---
title: "Robustness Checks for Acs (2025): Mapping the Political Contours of the Regulatory State: Dynamic Estimates of Agency Ideal Points"
subtitle: "PLSC 503: Homework 3"
author: "Jody Park"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
colorlinks: true
urlcolor: blue
linkcolor: blue
bibliography: references.bib
csl: apanodoi.csl
link-citations: true
header-includes:
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{placeins}   
  - \usepackage{setspace}
  - \usepackage{dcolumn}
  - \usepackage{booktabs}
  - \usepackage{siunitx}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{enumerate}
  - \usepackage[shortlabels]{enumitem}
  - \usepackage[hang,flushmargin]{footmisc}	
  - \usepackage{url}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{hyperref}
  - \hypersetup{
      colorlinks=true,
      linkcolor=blue,
      filecolor=magenta,      
      urlcolor=blue,
      citecolor=black,
    }
  - \usepackage[most]{tcolorbox}
  - \newtcolorbox{qbox}{enhanced,breakable,boxrule=0.5pt,arc=2mm,
    colback=gray!10,colframe=black!50,left=6pt,right=6pt,top=6pt,bottom=6pt}

---

```{r setup, include=FALSE}
options(knitr.table.format = "latex")
# clear all objects & free up memory if needed
rm(list = ls())

knitr::opts_chunk$set(
                      fig.width = 9, fig.height = 7, fig.retina = 3,
                      out.width = "100%",
                      cache = FALSE,
                      echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      hiline = TRUE
                      )
options(tinytex.verbose = TRUE)
knitr::opts_knit$set(allow_duplicate_labels = TRUE)

# set root directory to the project folder
# make sure whether the project folder exists in the folder
knitr::opts_knit$set(root.dir = normalizePath("."))

# check working directory
# getwd()

# load libraries
library(tidyverse)
library(ggpubr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(tidyr)
library(haven)
library(readr)
library(caret)
library(modelr)
library(broom)
library(ggplot2)
library(plotly)
library(png)
library(stringr)
library(openxlsx)
library(Hmisc)
library(MASS) 
library(lfe) 
library(modelsummary) 
library(stargazer)
library(lubridate) 
library(scales) 
library(rdrobust) 
library(psych) 
library(nFactors)
library(here)
library(rstanarm)
library(loo)
library(xtable)

# resolve conflicts
library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::summarize)
conflicts_prefer(dplyr::summarise)
conflicts_prefer(dplyr::rename)
conflicts_prefer(dplyr::lag)
conflicts_prefer(psych::describe)
conflicts_prefer(dplyr::recode)

select <- dplyr::select
filter <- dplyr::filter
summarize <- dplyr::summarize
summarise <- dplyr::summarise
rename <- dplyr::rename
lag <- dplyr::lag
describe <- psych::describe
recode <- dplyr::recode


# Acs (2025) replication setup
options(scipen=100)
options(digits=2)

# setwd("acs-2025-apsr")

#### ------------ PREAMBLE SCRIPTS (OPTIONAL) -------------------
## rebuild CSV files used in main results
# install.packages("XML")
# source('code/preamble_oira.R', local = TRUE)
# source('code/preamble_ua.R', local = TRUE)
# source('code/preamble_merge.R', local = TRUE)

# Fazekas et al. (2023) replication setup
# load the authors' constants in the replication code
source("../fazekas-ferrali-wachs-2023-jpart/functions.R")
nRep <- 1e3
set.seed(503) 
pdf <- TRUE # set to TRUE to output manuscript figures to pdf
eps <- TRUE # set to TRUE to output manuscript figures to eps

# set base size for figures
theme_set(theme_minimal(base_size = 14))

```


# 1. Replication Setup

## Citations and Links

- **Original Paper**: ACS, A. (2025). Mapping the Political Contours of the Regulatory State: Dynamic Estimates of Agency Ideal Points. American Political Science Review, 1-18.

- **Replication Data**: [R replication codes and dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/C7S2SI) from @acsmapping2025 "Mapping the Political Contours of the Regulatory State: Dynamic Estimates of Agency Ideal Points (APSR)."

- **Paper for the Extension**: Fazekas, M., Ferrali, R., & Wachs, J. (2023). Agency Independence, Campaign Contributions, and Favoritism in US Federal Government Contracting. Journal of Public Administration Research and Theory, 33(2), 262-278.

- **Replication Data**: [R replication codes and dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/3U07EE) from @fazekasagency2023 "Agency Independence, Campaign Contributions, and Favoritism in US Federal Government Contracting (JPART)." 

- **Repository Links**:  

  - [Dropbox folder](https://www.dropbox.com/scl/fo/fd8lfsrgua0ffxo3jncql/ALRWnyH1AW9au-xT7ztn2wk?rlkey=dl6drpt0kibpaete5r9bb5mgw&st=d0h6i6bv&dl=0); current code and data for this paper are in the *acs-2025-apsr* subfolder.

  - [GitHub repository for this project](https://github.com/jodypark/plsc503_replication_extension_fazekas2023/tree/main/acs-2025-apsr)

\bigskip

## Article Introduction

@acsmapping2025 investigates the political ideology of US federal regulatory agencies by estimating dynamic ideal points for 69 administrative units across four presidential administrations (Clinton to Trump). The study uses data from the Unified Agenda on planned regulations and the president's discretionary review of those regulations from the Office of Information and Regulatory Affairs (OIRA) docket. 
By analyzing the ideological positions of regulatory agencies over time, the paper provides insights into how agency ideology shapes regulatory behavior and how it interacts with political dynamics. The core contribution of the paper is to offer a comprehensive mapping of the political contours of the regulatory state, shedding light on the ideological underpinnings of regulatory activity and its implications for policy outcomes.

I extend the analysis by examining how agency ideology moderates the effects of political party contributions on favoritism in federal contracting, using the dynamic ideal point estimates from @acsmapping2025. Merging this agency ideology data with @fazekasagency2023 allows a direct test of whether ideological alignment between the president and an agency amplifies the returns to political contributions in procurement.
For this homework, I conduct three robustness checks to evaluate how much confidence the matched procurement sample can support: (1) a bootstrap to assess small-sample uncertainty, (2) a leave-one-agency-out sensitivity analysis to identify influential cases, and (3) a power-style simulation to evaluate the likelihood of detecting true effects given the sample size and effect sizes observed in the matched panel. 
These checks help determine whether the procurement results are precise enough to support substantive claims about the role of agency ideology in shaping procurement outcomes.

\bigskip

## Data, Variables and Models

### Data

In @acsmapping2025, the author uses publicly available datasets:

1. **Unified Agenda data**: Includes information on the regulations that federal agencies plan to implement, including their content, timing, and regulatory impact (a semi-annual catalog, indicating pending or completed).

2. **The Office of Information and Regulatory Affairs (OIRA) docket**: Lists the regulations reviewed. For each regulation, the docket includes information on the agency responsible, the date of review, and the outcome of the review (e.g., approved, disapproved, or returned for revision).

3. **Federal Register data**: Contains the official publication of federal regulations and related documents. [Federal Register API](https://www.federalregister.gov/developers/documentation/api/v1)


@acsmapping2025 uses the eight-digit regulation identifier number (the RIN) to merge the Unified Agenda data with the OIRA docket data.

- The analysis focuses on 69 administrative units (e.g., EPA, FDA, etc.) from 2001 to 2020, covering four presidential administrations (Clinton, Bush, Obama, Trump). 

  - The author excludes regulatory agencies that "did not issue a regulation in each administration or issued fewer than 40 regulations during the entire period (averaging less than 10 per administration)" [@acsmapping2025].
  
  - This exclusion criterion ensures that the analysis focuses on regulatory agencies with sufficient regulatory activity to estimate meaningful ideal points and examine their dynamics over time.
  

### Variables

1. **Dependent Variable**:

- (1) `Review` (`Y_i`) (for the dynamic measurement model): Binary indicator for whether a planned rule in the Unified Agenda was reviewed by OIRA (`1` = reviewed at least once in the OIRA docket; `0` = not reviewed).

- (2) `Regulatory Activity_{at}` (for regulatory productivity model): Agency-level regulatory output in administration `t`, measured using Federal Register final-rule page length (average per year within administration, with alternative coding rules reported in Table 2 and Supplementary Appendix).


2. **Independent Variable**: 

- `President-Agency ideological distance`: `|X_p - X_{at}|`, where `X_p` is the president's fixed ideal point (DW-NOMINATE anchors) and `X_{at}` is the agency ideal point in administration `t`.

3. **Model Components**:

- `Rule significance` (`S_i`): Binary indicator from Unified Agenda priority categories (`1` = economically significant/significant; `0` = otherwise), plus interaction with spatial distance in the measurement model.

- *Model structure*: `Agency intercepts` (`a_a`) and `dynamic structure`: Agency-specific baseline review propensity and a random-walk evolution of agency ideal points across administrations (Clinton, Bush, Obama, Trump), estimated in Stan (Supplementary Appendix B).


\newpage


### Models

For this homework, I will focus on *Model 2* (Regulatory Productivity) and *H3*, which use a negative binomial model to examine the relationship between President-Agency ideological distance and regulatory output. 

- **Model 1 (Dynamic Measurement Model / IRT-Probit)**: The core model estimates agency ideal points from OIRA review decisions using a Bayesian probit IRT structure with president-agency distance, significance terms, agency intercepts, and administration-specific agency ideal points that evolve via a random walk. The Stan specification and priors are provided in Supplementary Appendix B.

- **Model 2 (Validation Model: Regulatory Productivity)**: The paper then estimates a *negative binomial model* (Equation 3 in the original paper) where regulatory activity is regressed on President-Agency Distance with agency fixed effects and multiple outcome constructions (Table 2), testing whether ideologically closer agencies are more productive.

$$
\text{Regulatory Activity}_{at} \sim \text{NegBin}(\mu_{at}, \theta)
$$
$$
\ln(\mu_{at}) = \pi_a + \beta \cdot \text{PresidentAgencyDistance}_{at}
\tag{3}
$$


**Tested Hypotheses in the Original Paper**:

- **H1 (Review-distance expectation)**: OIRA review probability increases as ideological distance between the president and an agency increases.

- **H2 (Dynamic ideology expectation)**: Agency ideology is not static; agency ideal points shift across presidential administrations.

- **H3 (Productivity-distance expectation)**: Regulatory productivity is lower when President-Agency ideological distance is larger (negative association in Equation 3 / Table 2).


\newpage


# 2. Replication Accuracy

## Figure 4. Conservative Trends in Agency Ideal Points

```{r f4repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, floating = FALSE, fig.pos="H", fig.align="center", fig.cap = "Conservative Trends in Agency Ideal Points", out.width = "100%"}

# use the author's code for figure 4: ---

####  FIGURE 4 (conservative trends)
x <- read.csv(file = 'data/dynamic_estimates.csv')
x <- subset(x, x$agency != 'BENCHMARK')

## average ideal points by admin
##  - statistics used in section 'Temporal Dynamics of Agency Ideal Points'
admin_means <- tapply(x$est, x$admin, mean)
## weighted.avg <- function(var, se) sum(var * (1/se^2)) / sum(1/se^2)
## tapply(x$est, x$admin, weighted.avg, x$se)
## (.04 - (-.11)) / abs(-.11)

z <- split(x, x$agency.id)
m <- lapply(z, function(k) {
    k$est.max  <- max(k$est)
    k$est.min  <- min(k$est)
    k$admin.max <- k$admin[k$est == k$est.max]
    k$admin.min <- k$admin[k$est == k$est.min]
    k$dif <- sqrt(abs(k$est.max - k$est.min))
    return(k)
})
r <- do.call('rbind', m)
rownames(r) <- 1:nrow(r)
r <- r[!duplicated(r$agency.id), c('agency.id', 'est.max', 'est.min', 'admin.max', 'admin.min', 'dif')]
r <- r[order(r$dif, decreasing = TRUE),]
rownames(r) <- 1:nrow(r)

## collect biggest shifters (top 10)
a <- r$agency.id[1:10]
y <- x[x$agency.id %in% a,]
## identify 7 conservative shifts from top 10
a <- c('AID', 'TREAS-FINCEN', 'DOD-DARC', 'OPM', 'DOD-DODOASHA', 'VA', 'EPA-AR')

y$administration <- factor(y$time.period,
                              levels = c(1, 2, 3, 4),
                              labels = c("Clinton", "Bush", "Obama", "Trump"))
y <- y[y$agency.id %in% a,]

## recode names
y$group[y$agency.id == 'OPM' | y$agency.id == 'EPA-AR' |
        y$agency.id == 'AID'] <- 'Shift During \n Trump Administration'
y$group[y$agency.id == 'VA' | y$agency.id == 'DOD-DODOASHA'] <- 'Gradual Trend \n Liberal Agencies'
y$group[y$agency.id == 'TREAS-FINCEN'
        | y$agency.id == 'DOD-DARC'] <- 'Gradual Trend \n Conserv. Agencies'
y$group <- factor(y$group,
                  levels = c('Gradual Trend \n Liberal Agencies',
                             'Shift During \n Trump Administration',
                             'Gradual Trend \n Conserv. Agencies'))

y$agency.name <- y$agency.id
y$agency.name[y$agency.id == 'EPA-AR'] <- 'EPA (Office of Air)'
y$agency.name[y$agency.id == 'DOD-DODOASHA'] <- 'Health Affairs (DOD-OASHA)'
y$agency.name[y$agency.id == 'OPM'] <- 'Personnel Management (OPM)'
y$agency.name[y$agency.id == 'VA'] <- 'Veterans Affairs (VA)'
y$agency.name[y$agency.id == 'AID'] <- 'International Development (AID)'
y$agency.name[y$agency.id == 'DOD-DARC'] <- 'Defense Acquisition (DOD-DARC)'
y$agency.name[y$agency.id == 'TREAS-FINCEN'] <- 'Financial Crimes (TREAS-FINCEN)'

f4rep <- ggplot(y, aes(x = administration, y = est, linetype = agency.name, shape = agency.name, color = agency.name, group = agency.name)) +
  geom_line() +
   theme(plot.margin = unit(c(0.1, 0.5, 0.5, 0.5), "cm")) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "red") +
    facet_wrap(~ group) +
  scale_linetype_manual(values = 1:7) +
  scale_shape_manual(values = 1:7) +
    scale_color_manual(values = c("salmon", "salmon", "darkgreen", "salmon",
                                  "steelblue", "darkgreen", "steelblue")) +
  labs(title = "",
       x = "",
       y = "",
       linetype = "Agency",
       shape = "Agency",
       color = "Agency") +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major = element_line(linewidth = 0.5, colour = "grey"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.text = element_text(size = 8),
	        axis.text.x = element_text(angle = 45, hjust = 1)) +
	    guides(linetype = guide_legend(nrow = 4, title = NULL),
	           shape = guide_legend(nrow = 4, title = NULL),
	           color = guide_legend(nrow = 4, title = NULL))

print(f4rep)

# optional file export (keep the author's code)
dir.create("results", showWarnings = FALSE, recursive = TRUE)
dir.create("results/figure", showWarnings = FALSE, recursive = TRUE)
ggsave("results/figure4.pdf", plot = f4rep, width = 10, height = 7, units = "in")
ggsave("results/figure/f4repacs.png", plot = f4rep, width = 10, height = 7, dpi = 300)

```

Figure \@ref(fig:f4repacs) replicates the original pattern that several agencies shift in a conservative direction over administrations, with the largest movement concentrated in a small subset of agencies. This visual replication is consistent with @acsmapping2025's Figure 4.

\FloatBarrier

\newpage


## Table 1. Agencies with Largest Shifts in the Conservative Direction

```{r t1repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Agencies with Largest Shifts in the Conservative Direction", out.width = "100%"}

# use the author's code for table 1: ---

####  TABLE 1 (ideal point changes - follows from Figure 4)

# original code:
# ## use 'y' object from Figure 4
# 
# # Compute min and max by agency
# min.vals <- tapply(y$est, y$agency.id, min)
# max.vals <- tapply(y$est, y$agency.id, max)
# 
# # Combine into a data frame
# tab <- data.frame(
#   Agency = names(min.vals),
#   Min = as.numeric(min.vals),
#   Max = as.numeric(max.vals),
#   row.names = NULL
# )
# 
# # Create LaTeX table (NOTE: table is formatted differently in the paper)
# print(
#   xtable(tab,
#          align = c("l", "l", "r", "r")),
#   include.rownames = FALSE,
#   file = 'results/table1.tex')
# 


# use the original workflow object from Figure 4
stopifnot(exists("y"))

# Compute min and max by agency (original code logic)
min.vals <- tapply(y$est, y$agency.id, min)
max.vals <- tapply(y$est, y$agency.id, max)
tab <- data.frame(
  agency.id = names(min.vals),
  Min = as.numeric(min.vals),
  Max = as.numeric(max.vals),
  row.names = NULL
)

# paper order + labels/drivers
table1_labels <- tribble(
  ~agency.id,       ~Agency,                                  ~`Plausible Driver of Change`,                                ~ord,
  "OPM",            "Personnel Management (OPM)",             "\\makecell[l]{\\textit{Politicization:}\\\\Trump administration}",           1,
  "EPA-AR",         "EPA's Office of Air (EPA-AIR)",         "Trump administration",                                        2,
  "AID",            "Agency for Intl. Development (USAID)",  "Trump administration",                                        3,
  "VA",             "Veterans Affairs (VA)",                 "\\makecell[l]{\\textit{New policy conflict:}\\\\VA MISSION Act of 2018, wars}", 5,
  "DOD-DARC",       "Defense Acquisition (DOD-DARC)",        "Iraq/Afghan wars",                                            6,
  "TREAS-FINCEN",   "Financial Crimes (TREAS-FINCEN)",       "Patriot Act of 2001, civil liberties",                        7,
  "DOD-DODOASHA",   "Health Affairs, Def. (DOD-OASHA)",      "Iraq/Afghan wars",                                            8
)

table1_body <- table1_labels %>%
  left_join(tab, by = "agency.id") %>%
  mutate(
    Min = sprintf("%.2f", Min),
    Max = sprintf("%.2f", Max)
  ) %>%
  dplyr::select(ord, Agency, Min, Max, `Plausible Driver of Change`)

# spacer row between the two groups
table1_body <- bind_rows(
  table1_body,
  tibble(
    ord = 4,
    Agency = " ",
    Min = " ",
    Max = " ",
    `Plausible Driver of Change` = " "
  )
) %>%
  arrange(ord) %>%
  dplyr::select(-ord)

t1rep <- knitr::kable(
  table1_body,
  format = "latex",
  booktabs = TRUE,
  caption = "Agencies with Largest Shifts in the Conservative Direction",
  col.names = c("Agency", "Min", "Max", "Plausible Driver of Change"),
  align = c("l", "r", "r", "l"),
  escape = FALSE,
  linesep = ""
) %>%
  add_header_above(c(" " = 1, "Ideal Points" = 2, " " = 1), escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position")
  )

t1rep
t1rep %>% save_kable("results/table/t1repacs.tex")

```

Table \@ref(tab:t1repacs) reproduces the agency-level minima and maxima from the Figure 4 subset and highlights the same conservative movers. The ordering and substantive interpretation follow the original table.



\newpage


## Table 2. Ideological Distance and Regulatory Productivity (Negative Binomial)

```{r t2repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", out.width = "100%"}

# use the author's code for table 2: ---

####  TABLE 2 (productivity in number of rules)

# library(stargazer)
# library(lubridate)
# library(MASS)

v <- read.csv(file = 'data/estimation_data.csv')

v$agency.id[v$agency.abb == v$bureau.abb] <- v$agency.abb[v$agency.abb == v$bureau.abb]
v$agency.code <- substr(v$rin, 1, 4)
v <- v[!duplicated(v$agency.code), c('agency.id', 'agency.code')]
rownames(v) <- 1:nrow(v)
v$agency.code <- sprintf("%04s", v$agency.code)

y <- read.csv(file = 'data/dynamic_estimates.csv')

v$agency.id[!v$agency.id %in% y$agency.id] <- gsub('(.*)-.*', '\\1', v$agency.id[!v$agency.id %in% y$agency.id])

y <- merge(y, v, by = 'agency.id')
n_agency_admin <- nrow(y) # 276

## import OIRA data to get RINs of reviewed rules
load('data/oira_data/oira_data.RData')
g <- oira.data
colnames(g) <- c('agency.code', 'rin', 'rule', 'stage', 'econ.sig', 'date.received',
                 'legal.dline', 'date.completed', 'decision', 'date.published')
## recode variables
g$returned <- ifelse(g$decision == 'Returned for Reconsideration' |
                     g$decision == 'Withdrawn', 1, 0)
g$accepted <- ifelse(g$decision == 'Consistent without Change', 1, 0)
g$review.duration <- as.numeric(as.Date(g$date.completed) - as.Date(g$date.received))

r <- subset(g, date.received > '1995-01-01' & date.received < '2021-01-19')
r <- subset(r, stage == 'Proposed Rule')
## drop odd cases
d <- c('Exempt from Executive Order', 'Improperly Submitted', 'Emergency',
       'Statutory or Judicial Deadline')
r <- subset(r, !decision %in% d)
r$accepted <- ifelse(r$decision == 'Consistent without Change', 1, 0)
r$total <- rep(1, nrow(r))
r$admin <- rep('Clinton', nrow(r))
r$admin[as.Date(r$date.received) > '2001-01-19'] <- 'Bush'
r$admin[as.Date(r$date.received) > '2009-01-19'] <- 'Obama'
r$admin[as.Date(r$date.received) > '2016-01-19'] <- 'Trump'

r1 <- aggregate(cbind(returned, accepted, total) ~ agency.code + admin, data = r, sum, na.rm = TRUE)
r2 <- aggregate(review.duration ~ agency.code + admin, data = r, mean, na.rm = TRUE)
r1 <- merge(r1, r2, by = c('agency.code', 'admin'), all.x = TRUE)

v.expand <- expand.grid(agency.id = v$agency.id,
                    admin = c('Clinton', 'Bush', 'Obama', 'Trump'))
v.expand <- merge(v, v.expand, by =  'agency.id')
r1 <- merge(v.expand, r1, by = c('agency.code', 'admin'), all.x = T)
r1[is.na(r1)] <- 0 ## agencies with no reviews get count of 0

r3 <- aggregate(review.duration ~ agency.code + admin,
                data = subset(r, econ.sig == 'Yes'), mean, na.rm = TRUE)
colnames(r3)[3] <- 'review.dur.econ'
r <- merge(r1, r3, by = c('agency.code', 'admin'), all.x = TRUE)

## federal register data on final rules
x <- read.csv(file = 'data/final_rules_from_api.csv')

## fix date and admin
x$date <- mdy(x$publication_date)
x$date <- format(x$date, "%Y-%m-%d")
x$admin <- rep('Clinton', nrow(x))
x$admin[as.Date(x$date) > '2001-01-19'] <- 'Bush'
x$admin[as.Date(x$date) > '2009-01-19'] <- 'Obama'
x$admin[as.Date(x$date) > '2017-01-19'] <- 'Trump'
x$year <- substr(x$date, 1, 4)
x <- subset(x, date < '2021-01-20')

sig_table_raw <- table(x$significant)
x$sig <- x$significant
x$sig[x$significant == 'true'] <- 1
x$sig[x$significant == 'false'] <- 0
sig_table_binary <- table(x$sig)

x$rin <- gsub('^(.*?);.*', '\\1', x$regulation_id_numbers)

x$agency.code <- substr(x$rin, 1, 4)
x <- subset(x, !is.na(x$agency.code))
## counter for the number of rules
x$counter <- rep(1, nrow(x))

## fix action
n_action_raw <- length(unique(x$action))
x$action.new <- x$action
x$action.new <- gsub('ACTION:', '', x$action.new)
x$action.new <- trimws(x$action.new)
x$action.new <- gsub('  +', ' ', x$action.new)
x$action.new <- gsub('\\.$', '', x$action.new)
x$action.new <- tolower(x$action.new)
n_action_clean <- length(unique(x$action.new))

## categories to isolate final rules
x$action.new[grep('rescission|removal|commentary|temporary|technical amend|notice of enforcement|bridge|catch|final special conditions|withdraw|quota|immediate final rule|notice of availability|closure|clarification|sunset review|retention limits|delay|information collection|deferral|petition|correct|clarification|conditional|confirmation|compliance date|compliance deadline|extension|extend|expiration|change of|approval of|request for|court order|adjustments|modification|stay|interim|direct|information collection|advisory opinion|^notice$|interpret|notification of|notice of.*date|small entity compliance guide|meeting|guid', x$action.new)] <- 'direct.final.etc'
n_action_filtered <- length(unique(x$action.new))

## ASSUMPTION: what counts as a final rule
## drop.direct.rules = 0
drop.direct.rules = 1
if(drop.direct.rules == 0) {
    x$final.rule <- rep(1, nrow(x))
} else {
    x$final.rule <- rep(0, nrow(x))
    x$final.rule[grep('final rule|final reg|final order|final policy', x$action.new)] <- 1
}
final_rule_table <- table(x$final.rule)

## search for deregulatory action
## keywords associated with deregulation
keywords <- c("remove", "eliminate", "repeal", "reduce", "lower", "decrease",
              "simplify", "streamline", "ease", "exempt", "exclusion", "waive",
              "deregulate", "flexibility", "burden reduction", ## "flexibility act",
              "relief", "less burdensome", "sunset", "expiration", "reform", # "modernize"
              "cost savings", "efficiency", "remove barriers",
              "rescind", "decreased oversight", "less supervision", "voluntary compliance")
# Combine the keywords into a single regex pattern
pattern <- paste(keywords, collapse = "|")
x$dereg <- rep(0, nrow(x))
x$dereg[grep(pattern, x$abstract, ignore.case = TRUE)] <- 1
x$counter.regulatory <- x$counter
x$counter.regulatory[x$dereg == 1] <- 0
x$final.regulatory <- x$final.rule
x$final.regulatory[x$dereg == 1] <- 0

x$n.pages <- x$page_length
x$n.pages.regulatory <- x$page_length
x$n.pages.regulatory[x$dereg == 1] <- 0
x$n.pages.final.rule <- x$n.pages
x$n.pages.final.rule[x$final.rule == 0] <- 0

x$n.pages.final.rule.regulatory <- x$n.pages
x$n.pages.final.rule.regulatory[x$final.rule == 0] <- 0
x$n.pages.final.rule.regulatory[x$dereg == 1] <- 0

## significant
x$n.pages.final.rule.sig <- x$n.pages
x$n.pages.final.rule.sig[x$final.rule == 0] <- 0
x$n.pages.final.rule.sig[x$sig != 1] <- 0

## significant & regulatory
x$n.pages.final.rule.sig.reg <- x$n.pages
x$n.pages.final.rule.sig.reg[x$final.rule == 0] <- 0
x$n.pages.final.rule.sig.reg[x$sig != 1] <- 0
x$n.pages.final.rule.sig.reg[x$dereg == 1] <- 0
## x$n.pages.final.rule.sig[x$dereg == 1] <- 0

g <- g[g$stage == 'Proposed Rule',] # focus on reivew of proposal
x$n.pages.no.review <- x$n.pages  ## this version does not drop when x$final.rule == 0
x$n.pages.no.review[x$rin %in% unique(g$rin)] <- 0 # drop pages from reviewed rules
x$n.pages.no.review.final <- x$n.pages.no.review
x$n.pages.no.review.final[x$final.rule == 0] <- 0 # drop non-final rules
x$n.pages.no.review.final.dereg <- x$n.pages.no.review.final
x$n.pages.no.review.final.dereg[x$dereg == 1] <- 0 # drop de-regulatory rules

robust = 0
if(robust == 1) {
    x <- x[x$year != 2001 &
           x$year != 2009 &
           x$year != 2017,]
}

z <- aggregate(cbind(counter, counter.regulatory, final.rule, n.pages,
                     n.pages.final.rule, n.pages.final.rule.regulatory,
                     n.pages.regulatory, final.regulatory,
                     n.pages.final.rule.sig, n.pages.final.rule.sig.reg,
                     n.pages.no.review, n.pages.no.review.final,
                     n.pages.no.review.final.dereg) ~
               agency.code + admin, data = x, sum)
w <- aggregate(dereg ~ agency.code + admin, data = x, mean)
z <- merge(z, w, by = c('agency.code', 'admin'))

## y$agency.code %in% z$agency.code # all present

y <- merge(y, z, by = c('agency.code', 'admin'), all.x = T)
y[is.na(y)] <- 0

y$pres.ideal <- rep(.693, nrow(y))
y$pres.ideal[y$admin == 'Clinton'] <- -.438
y$pres.ideal[y$admin == 'Obama'] <- -.358
y$pres.ideal[y$admin == 'Trump'] <- .403
y$distance <- abs(y$pres.ideal - y$est)

year.adjust <- function(var, admin) {
    var[admin == 'Clinton'] <- var[admin == 'Clinton'] / 6 # (1995 - 2000)
    var[admin == 'Obama'] <- var[admin == 'Obama'] / 8
    var[admin == 'Bush'] <- var[admin == 'Bush'] / 8
    var[admin == 'Trump'] <- var[admin == 'Trump'] / 4
    return(var)
}

if(robust == 1) {
    year.adjust <- function(var, admin) {
        var[admin == 'Clinton'] <- var[admin == 'Clinton'] / 6 # (1995 - 2000)
        var[admin == 'Obama'] <- var[admin == 'Obama'] / 7
        var[admin == 'Bush'] <- var[admin == 'Bush'] / 7
        var[admin == 'Trump'] <- var[admin == 'Trump'] / 3
        return(var)
    }
}

y$counter.adj <- year.adjust(y$counter, y$admin)
y$counter.regulatory.adj <- year.adjust(y$counter.regulatory, y$admin)
y$final.rule.adj <- year.adjust(y$final.rule, y$admin)
y$final.regulatory.adj <- year.adjust(y$final.regulatory, y$admin)
y$n.pages.adj <- year.adjust(y$n.pages, y$admin)
y$n.pages.regulatory.adj <- year.adjust(y$n.pages.regulatory, y$admin)

y$n.pages.final.rule.adj <- year.adjust(y$n.pages.final.rule, y$admin)
y$n.pages.final.rule.regulatory.adj <- year.adjust(y$n.pages.final.rule.regulatory, y$admin)

y$n.pages.final.rule.sig.adj <- year.adjust(y$n.pages.final.rule.sig, y$admin)
y$n.pages.final.rule.sig.reg.adj <- year.adjust(y$n.pages.final.rule.sig.reg, y$admin)


y$n.pages.no.review.adj <- year.adjust(y$n.pages.no.review, y$admin)
y$n.pages.no.review.final.adj <- year.adjust(y$n.pages.no.review.final, y$admin)
y$n.pages.no.review.final.dereg.adj <- year.adjust(y$n.pages.no.review.final.dereg, y$admin)


n_obs_y <- nrow(y)
a <- aggregate(est ~ agency.id, data = y, mean)
colnames(a)[2] <- 'est.avg'
y <- merge(y, a, by = 'agency.id')
y$liberal.agency <- rep(0, nrow(y))
y$liberal.agency[y$est.avg < 0] <- 1

# keep a dedicated ACS panel object for later extension chunks
y_acs <- y

agency.fe <- c('Agency Fixed Effects', rep('$\\checkmark$', 5))

n <- c('President-Agency Distance')

# glm.nb() models for the five outcome constructions in Table 2, with distance and agency fixed effects, weighted by the inverse variance of each agency ideal point (1/se^2)
m1 <- glm.nb(as.integer(n.pages.final.rule.adj) ~ distance + agency.id, data = y, weights = 1/se^2)

m2 <- glm.nb(as.integer(n.pages.final.rule.sig.adj) ~ distance + agency.id, data = y, weights = 1/se^2)
m3 <- glm.nb(as.integer(n.pages.final.rule.sig.reg.adj) ~ distance + agency.id, data = y, weights = 1/se^2)

m4 <- glm.nb(as.integer(n.pages.no.review.final.adj) ~ distance + agency.id, data = y, weights = 1/se^2)

m5 <- glm.nb(as.integer(n.pages.no.review.final.dereg.adj) ~ distance + agency.id, data = y,
             weights = 1/se^2)

## exp(coef(m1)['distance']) - 1
## exp(coef(m2)['distance']) - 1
## exp(coef(m3)['distance']) - 1

options(digits=3)
dir.create("results", showWarnings = FALSE, recursive = TRUE)

# original code
# stargazer(m1, m2, m3, m4, m5,
#           ## type = 'text',
#           out = 'results/table2.tex',
#           omit = 'Constant|agency',
#           label = 'tbl:reg',
#           digits = 3,
#           title = 'Ideological Distance and Regulatory Productivity',
#           dep.var.labels = c('Formal', 'formal and sig', 'form, sig and Reg.', 'formal', 'formal and reg'),
#           star.cutoffs = c(.1, .05, .01),
#           star.char = c('^\\cdot', '*', '**'),
#           covariate.labels = n,
#           model.names = F,
#           add.lines = list(agency.fe),
#           no.space = TRUE,
#           font.size = 'small',
#           omit.stat=c("f", 'ser', 'adj.rsq', 'aic', 'bic', 'theta'))

# format coefficient and standard error for the distance row
fmt_t2 <- function(model) {
  s <- coef(summary(model))
  est <- unname(s["distance", "Estimate"])
  se  <- unname(s["distance", "Std. Error"])
  p   <- unname(s["distance", "Pr(>|z|)"])
  list(
    est = ifelse(
      !is.na(p) && p < 0.05,
      sprintf("$%0.3f^{*}$", est),
      sprintf("$%0.3f$", est)
    ),
    se  = sprintf("(%0.3f)", se)
  )
}

t2_stats <- lapply(list(m1, m2, m3, m4, m5), fmt_t2)
t2_est <- vapply(t2_stats, `[[`, character(1), "est")
t2_se  <- vapply(t2_stats, `[[`, character(1), "se")
t2_n   <- vapply(list(m1, m2, m3, m4, m5), nobs, numeric(1))

# keep the format like table 2 in the paper
t2rep <- c(
  "\\begin{table}[H]",
  "\\centering",
  "\\caption{Ideological Distance and Regulatory Productivity (Negative Binomial)}",
  "\\label{tab:t2repacs}",
  "\\small",
  "\\begin{tabular}{lccccc}",
  "\\hline\\hline",
  " & \\multicolumn{5}{c}{Dependent variable: Regulatory activity (final rule page length)} \\\\",
  "\\cline{2-6}",
  " & (1) & (2) & (3) & (4) & (5) \\\\",
  paste0(
    " & Formal & \\makecell[c]{Formal\\\\significant} & ",
    "\\makecell[c]{Formal significant\\\\regulatory} & ",
    "\\makecell[c]{Formal\\\\(excl. OIRA)} & ",
    "\\makecell[c]{Formal regulatory\\\\(excl. OIRA)} \\\\"
  ),
  "\\hline",
  paste0("Pres.--Agency Distance & ", paste(t2_est, collapse = " & "), " \\\\"),
  paste0(" & ", paste(t2_se, collapse = " & "), " \\\\"),
  "\\hline",
  "Agency Fixed Effects & $\\checkmark$ & $\\checkmark$ & $\\checkmark$ & $\\checkmark$ & $\\checkmark$ \\\\",
  paste0("Observations & ", paste(format(t2_n, trim = TRUE, scientific = FALSE), collapse = " & "), " \\\\"),
  "\\hline",
  paste0(
    "\\multicolumn{6}{l}{\\parbox{0.95\\textwidth}{$^{*}p < 0.05$. ",
    "Standard errors in parentheses. Models 4-5 exclude rules reviewed by OIRA at the proposed rule stage. ",
    "Models 3 and 5 also exclude deregulatory rules. Weights equal to the inverse variance of each agency ideal point are included.}} \\\\"
  ),
  "\\hline\\hline",
  "\\end{tabular}",
  "\\end{table}"
)

dir.create("results/table", showWarnings = FALSE, recursive = TRUE)
writeLines(t2rep, "results/table/t2repacs.tex")
cat(t2rep, sep = "\n")

```

Table \@ref(tab:t2repacs) replicates the negative-binomial validation result: larger President-Agency ideological distance is associated with lower regulatory productivity across outcome definitions. The sign and substantive magnitude patterns are aligned with the original paper.


\newpage


## Table A1. Agency Names and Ideal Point Estimates

```{r ta1repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Agency Names and Ideal Point Estimates", out.width = "100%"}

# use the author's code for table a1: ---

####   - TABLE A1

x <- read.csv(file = 'data/estimation_data.csv')
x <- subset(x, x$agency.id != 'BENCHMARK')

# get dept names
a <- read.csv(file = 'data/ua_data/agencycodes.csv')
a$agency.id <- paste(a$agency.abb, a$bureau.abb, sep = '-')
a <- subset(a, select = c(agency.id, agency, bureau, agency.abb, bureau.abb) )
a <- a[!duplicated(a),]
x <- merge(x, a, by = c('agency.id'), all.x = T)

x$agency.id.new <- x$agency.id
x$agency.id.new[x$agency.abb == x$bureau.abb] <- x$agency.abb[x$agency.abb == x$bureau.abb]

## create agency name table (APPENDIX TABLE 1)
x$agency.parent.new <- x$agency
x$agency.parent.new[x$agency.abb == x$bureau.abb] <- ''


y <- subset(x, select = c(agency.id.new, admin, sig, review))

#library(dplyr)
#library(tidyr)
#library(xtable)

collapsed_data <- y %>%
  group_by(admin, agency.id.new) %>%
  summarise(
    sig_mean = mean(sig, na.rm = TRUE),
    review_mean = mean(review, na.rm = TRUE),
    counter = n()
  ) %>%
  ungroup()

# convert to wide format
collapsed_pivot <- collapsed_data %>%
  pivot_wider(
    names_from = admin,
    values_from = c(sig_mean, review_mean, counter),
    names_glue = "{admin}_{.value}"
  )

z <- as.data.frame(collapsed_pivot)

# order the columns
ordered_columns <- c(
  "agency.id.new",
  "clinton_sig_mean", "clinton_review_mean", "clinton_counter",
  "bush43_sig_mean", "bush43_review_mean", "bush43_counter",
  "obama_sig_mean", "obama_review_mean", "obama_counter",
  "trump_sig_mean", "trump_review_mean", "trump_counter"
)

# reorder the dataframe columns
z <- z[, ordered_columns]

rownames(z) <- NULL
print(xtable(z),
      include.rownames = F,
      file = 'results/tableA1.tex')


# table a1 replication
# make display names latex-safe and readable in the knitted table
z_display <- z
colnames(z_display) <- c(
  "agency id",
  "clinton sig mean", "clinton review mean", "clinton n",
  "bush43 sig mean", "bush43 review mean", "bush43 n",
  "obama sig mean", "obama review mean", "obama n",
  "trump sig mean", "trump review mean", "trump n"
)

ta1rep_obj <- knitr::kable(
  z_display,
  format = "latex",
  booktabs = TRUE,
  caption = "Agency Names and Ideal Point Estimates",
  align = c("l", rep("c", ncol(z_display) - 1)),
  escape = TRUE
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  )

ta1rep_obj
ta1rep_obj %>% save_kable("results/table/ta1repacs.tex")
cat("\n\\FloatBarrier\n")

```

Table \@ref(tab:ta1repacs) reproduces the appendix-style agency-by-administration summaries of significance rates, review rates, and counts. These summaries support the data pipeline used to estimate the main replication models.


\newpage



# 3. Methodological Extension

## Why this method is necessary

Section 2 shows that I can reproduce the main ACS validation results with the original agency-administration panel. My [HW2 extension](https://www.dropbox.com/scl/fi/uwsy111lt5x67ssn24vnr/park_robustness_check_acs-2025_503_260419.pdf?rlkey=8nb7r5caignkz9j09i8cj4yv4&st=1z9u4b4n&dl=0), however, depends on a much smaller matched ACS-Fazekas panel. The ACS replication in Section 2 uses an agency-administration panel with `N = 276`, which is large enough to estimate the negative binomial validation model in @acsmapping2025 with relatively stable coefficients. 
Once I merge ACS ideological distance with the contracting data, the effective sample size falls to 48 agency-administration observations in the main procurement models, 42 in the ordered logit, and 22 in the agency-level follow-up. The matched procurement panel contains 25 distinct agencies, and only 22 remain once the agency-independence categories needed for the ordered-logit checks are imposed. 

That drop in sample size creates a different inferential problem than the one emphasized in the original ACS article. The concern is no longer whether the model form is correct. Instead, the concern is whether coefficient instability reflects substantive heterogeneity or simply limited information. With a sample this small, null or mixed findings may be low-power results rather than strong evidence against the extension. 

## How the method improves the paper

For HW3, I keep the HW2 extension design and reframe the robustness question around sample size and power. I use three checks. First, I bootstrap the key `distance` coefficients in the four procurement models to quantify how much small-sample uncertainty widens the confidence intervals beyond the model-based standard errors. Second, I conduct a leave-one-agency-out sensitivity analysis to determine whether the results are driven by a handful of matched agencies. Third, I run a power-style simulation that projects rejection rates as the effective sample size increases from the current matched panel toward the size of the original ACS replication sample. 

These checks do not replace the extension models. They evaluate whether the extension results are fragile because the matched panel is small. This is suitable for HW3 because the sample size problem is already visible in the current write-up, and because the assignment specifically asks for a robustness check that addresses sample size and power. 

### Robustness Check 1: Small-Sample Uncertainty via Bootstrap

The first robustness check resamples agencies with replacement and refits the four procurement models from HW2. This is an agency bootstrap rather than a row bootstrap because the two administration rows for a given agency are not substantively independent. Resampling at the agency level preserves that within-agency structure while still quantifying how much the `distance` coefficient varies under repeated sampling from a small matched panel. 

If the bootstrap intervals are much wider than the asymptotic intervals from the original models, then the extension results are more fragile than the default standard errors suggest. If the sign remains stable and the bootstrap interval still excludes zero, then the small sample is less threatening to the substantive conclusion.

### Robustness Check 2: Leave-One-Agency-Out Sensitivity

The second robustness check drops one matched agency at a time and re-estimates the four procurement models. This directly tests whether the extension is driven by a small number of influential agencies. With only 25 matched agencies, a single omitted case can shift the point estimate in a meaningful way. 

The substantive question is: do the coefficients keep the same sign and roughly the same magnitude when each agency is omitted in turn? 
If the answer is yes, then the extension is not being carried by one or two special cases. If the sign flips or the coefficient collapses toward zero under many omissions, then the extension should be interpreted more cautiously.


### Robustness Check 3: Power-Style Simulation

The third robustness check asks whether the current matched panel is large enough to detect effects of the size observed in HW2. I treat the estimated coefficient and standard error from each procurement model as the starting point for a retrospective design analysis. Then I project what power would look like at the current extension sample size (`N = 48`), a doubled sample (`N = 96`), a four-times larger sample (`N = 192`), and the original ACS replication size (`N = 276`). 

I report power under two assumptions: the observed effect size from the HW2 extension, and a more conservative scenario in which the true effect is only half as large. This check follows the sample-size logic from class: if power remains low even under optimistic assumptions, then coefficient instability is more likely to reflect limited information than a well-estimated null effect.

## Explanatory Table

### Explanatory Table 4. Effective Sample Sizes Across ACS Replication and Extension Models

```{r robustpanelsetup, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Effective sample sizes across ACS replication and extension models", out.width = "100%"}

# make sure HW3 output folders exist
dir.create("results", showWarnings = FALSE, recursive = TRUE)
dir.create("results/table", showWarnings = FALSE, recursive = TRUE)
dir.create("results/figure", showWarnings = FALSE, recursive = TRUE)

# standardize agency names before matching
clean_agency_text <- function(x) {
  x %>%
    str_to_lower() %>%
    str_replace_all("&", " and ") %>%
    str_replace_all("\\badmin\\.?\\b", "administration") %>%
    str_replace_all("[^a-z0-9]+", " ") %>%
    str_squish()
}

# build ACS distance by agency and administration
acs_dynamic_ext <- read_csv("data/dynamic_estimates.csv", show_col_types = FALSE) %>%
  filter(agency.id != "BENCHMARK") %>%
  mutate(
    pres_ideal = case_when(
      admin == "Clinton" ~ -0.438,
      admin == "Bush" ~ 0.693,
      admin == "Obama" ~ -0.358,
      admin == "Trump" ~ 0.403,
      TRUE ~ NA_real_
    ),
    distance = abs(pres_ideal - est)
  ) %>%
  select(agency.id, admin, distance, est, se) %>%
  distinct()

# aggregate ACS-side productivity controls
acs_prod_ext <- read_csv("data/estimation_data.csv", show_col_types = FALSE) %>%
  mutate(
    sig = as.numeric(sig),
    review = as.numeric(review)
  ) %>%
  group_by(agency.id, admin) %>%
  summarize(
    n_rules = n(),
    n_sig_rules = sum(sig == 1, na.rm = TRUE),
    review_rate = mean(review, na.rm = TRUE),
    .groups = "drop"
  )

# prepare ACS agency names for matching
acs_name_ext <- read_csv("data/rcl_ideal_points_merged.csv", show_col_types = FALSE) %>%
  transmute(
    agency.id,
    agency_name_acs = coalesce(cleaned_agency, agency.y, agency.x)
  ) %>%
  filter(!is.na(agency_name_acs)) %>%
  distinct() %>%
  mutate(agency_name_clean = clean_agency_text(agency_name_acs)) %>%
  arrange(agency.id) %>%
  distinct(agency_name_clean, .keep_all = TRUE)

# recode Fazekas contracts to agency-admin outcomes
faz_ext <- read_csv("../fazekas-ferrali-wachs-2023-jpart/contracts.csv", show_col_types = FALSE) %>%
  mutate(
    donation_total = pmax(current_R_donor, 0) + pmax(current_D_donor, 0),
    donation_dummy = as.integer(donation_total > 0),
    admin = case_when(
      year <= 2008 ~ "Bush",
      year >= 2009 & year <= 2016 ~ "Obama",
      year < 2001 ~ "Clinton",
      TRUE ~ "Trump"
    ),
    agency_name_clean = clean_agency_text(agency_name)
  ) %>%
  filter(admin %in% c("Bush", "Obama"))

# merge the ACS and Fazekas pieces into one panel
ext_panel <- faz_ext %>%
  group_by(agency_name, agency_name_clean, admin) %>%
  summarize(
    n_contracts = n(),
    donor_contract_share = mean(donation_dummy, na.rm = TRUE),
    donor_contract_count = sum(donation_dummy, na.rm = TRUE),
    donation_amount = sum(donation_total, na.rm = TRUE),
    unique_firms = n_distinct(pduns),
    .groups = "drop"
  ) %>%
  left_join(
    acs_name_ext %>% select(agency_name_clean, agency.id),
    by = "agency_name_clean"
  ) %>%
  left_join(acs_dynamic_ext, by = c("agency.id", "admin")) %>%
  left_join(acs_prod_ext, by = c("agency.id", "admin")) %>%
  filter(!is.na(distance)) %>%
  mutate(
    n_contracts = pmax(n_contracts, 1),
    n_rules = replace_na(n_rules, 0),
    n_sig_rules = replace_na(n_sig_rules, 0),
    review_rate = replace_na(review_rate, 0)
  )

# recover agency independence categories from the Fazekas data
faz_cat_lookup <- read_csv(
  "../fazekas-ferrali-wachs-2023-jpart/contracts.csv",
  show_col_types = FALSE
) %>%
  mutate(agency_name_clean = clean_agency_text(agency_name)) %>%
  group_by(agency_name_clean) %>%
  summarize(agency_cat2 = first(agency_cat2), .groups = "drop") %>%
  filter(!is.na(agency_cat2))

# order categories from most to least independent
indep_levels <- c(
  "indep. commission & IRC",
  "indep. agency",
  "cabinet/exec. dep. and bureau",
  "cabinet/exec. dep. (not bureau)"
)

# merge categories for ordered logit variants
ext_panel_cat <- ext_panel %>%
  left_join(faz_cat_lookup, by = "agency_name_clean") %>%
  filter(!is.na(agency_cat2), !is.na(distance)) %>%
  mutate(
    agency_cat_ord = factor(
      agency_cat2,
      levels = indep_levels,
      ordered = TRUE
    )
  )

# collapse to one row per agency for the smallest follow-up model
ext_panel_agency <- ext_panel_cat %>%
  group_by(agency.id, agency_name_clean, agency_cat2) %>%
  summarize(
    mean_distance = mean(distance, na.rm = TRUE),
    n_admin = n(),
    .groups = "drop"
  ) %>%
  mutate(
    agency_cat_ord = factor(
      agency_cat2,
      levels = indep_levels,
      ordered = TRUE
    )
  )

# fit the four baseline procurement models from HW2
ext_mods <- list(
  "Donor contract share" = lm(
    donor_contract_share ~ distance + log1p(n_rules) + log1p(n_sig_rules) + review_rate + factor(admin),
    data = ext_panel
  ),
  "Donor contract count" = glm.nb(
    donor_contract_count ~ distance + log1p(n_rules) + log1p(n_sig_rules) + factor(admin) + offset(log(n_contracts)),
    data = ext_panel
  ),
  "Log donation amount" = lm(
    log1p(donation_amount) ~ distance + log1p(n_rules) + log1p(n_sig_rules) + review_rate + factor(admin),
    data = ext_panel
  ),
  "Total contracts" = glm.nb(
    n_contracts ~ distance + factor(admin),
    data = ext_panel
  )
)

# compact sample-size overview table
sample_size_overview <- tibble(
  Model = c(
    "ACS replication Table 2",
    "HW3 matched procurement panel",
    "Ordered logit extension",
    "Agency-level ordered logit"
  ),
  Unit = c(
    "Agency-administration",
    "Agency-administration",
    "Agency-administration",
    "Agency"
  ),
  `Num.Obs.` = c(
    n_agency_admin,
    nrow(ext_panel),
    nrow(ext_panel_cat),
    nrow(ext_panel_agency)
  ),
  `Distinct agencies` = c(
    n_distinct(y_acs$agency.id),
    n_distinct(ext_panel$agency.id),
    n_distinct(ext_panel_cat$agency.id),
    n_distinct(ext_panel_agency$agency.id)
  )
)

sample_kbl <- kable(
  sample_size_overview,
  format = "latex",
  booktabs = TRUE,
  caption = "Effective sample sizes across ACS replication and extension models",
  align = c("l", "l", "r", "r")
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position")
  )

sample_kbl
sample_kbl %>% save_kable("results/table/sample_size_overview_hw3.tex")

```

Table \@ref(tab:robustpanelsetup) shows the core inferential problem motivating HW3. The original ACS replication uses 276 agency-administration observations, but the matched procurement extension falls to 48 rows, and the ordered-logit follow-up models shrink further to 42 and 22. This large drop in effective sample size is the basis for the robustness checks below.

## Robustness Check 1. Small-Sample Uncertainty via Bootstrap

```{r robustboot, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Bootstrap uncertainty for procurement extension coefficients", out.width = "100%"}

stopifnot(exists("ext_panel"), exists("ext_mods"))

# agency bootstrap helper
resample_agencies <- function(data) {
  ids <- unique(data$agency.id)
  draw <- sample(ids, size = length(ids), replace = TRUE)
  map2_dfr(draw, seq_along(draw), ~ {
    data %>%
      filter(agency.id == .x) %>%
      mutate(boot_cluster = .y)
  })
}

# fit each baseline model on a resampled panel
fit_boot_models <- function(data) {
  list(
    share = tryCatch(
      lm(donor_contract_share ~ distance + log1p(n_rules) + log1p(n_sig_rules) + review_rate + factor(admin), data = data),
      error = function(e) NULL
    ),
    count = tryCatch(
      suppressWarnings(glm.nb(donor_contract_count ~ distance + log1p(n_rules) + log1p(n_sig_rules) + factor(admin) + offset(log(n_contracts)), data = data)),
      error = function(e) NULL
    ),
    amount = tryCatch(
      lm(log1p(donation_amount) ~ distance + log1p(n_rules) + log1p(n_sig_rules) + review_rate + factor(admin), data = data),
      error = function(e) NULL
    ),
    contracts = tryCatch(
      suppressWarnings(glm.nb(n_contracts ~ distance + factor(admin), data = data)),
      error = function(e) NULL
    )
  )
}

# collect the baseline estimates and asymptotic intervals
baseline_summary <- bind_rows(
  broom::tidy(ext_mods[[1]], conf.int = TRUE) %>% mutate(model = "Donor contract share"),
  broom::tidy(ext_mods[[2]], conf.int = TRUE) %>% mutate(model = "Donor contract count"),
  broom::tidy(ext_mods[[3]], conf.int = TRUE) %>% mutate(model = "Log donation amount"),
  broom::tidy(ext_mods[[4]], conf.int = TRUE) %>% mutate(model = "Total contracts")
) %>%
  filter(term == "distance") %>%
  transmute(
    model,
    estimate,
    std.error,
    asym_lo = conf.low,
    asym_hi = conf.high
  )

# bootstrap the distance coefficient by resampling agencies
set.seed(503)
boot_reps <- 199
boot_draws <- vector("list", boot_reps)

for (b in seq_len(boot_reps)) {
  boot_dat <- resample_agencies(ext_panel)
  boot_fit <- fit_boot_models(boot_dat)
  boot_draws[[b]] <- tibble(
    rep = b,
    model = c("Donor contract share", "Donor contract count", "Log donation amount", "Total contracts"),
    estimate = c(
      if (!is.null(boot_fit$share)) coef(boot_fit$share)[["distance"]] else NA_real_,
      if (!is.null(boot_fit$count)) coef(boot_fit$count)[["distance"]] else NA_real_,
      if (!is.null(boot_fit$amount)) coef(boot_fit$amount)[["distance"]] else NA_real_,
      if (!is.null(boot_fit$contracts)) coef(boot_fit$contracts)[["distance"]] else NA_real_
    )
  )
}

boot_draws <- bind_rows(boot_draws)

boot_summary_raw <- boot_draws %>%
  group_by(model) %>%
  summarize(
    boot_lo = quantile(estimate, 0.025, na.rm = TRUE),
    boot_hi = quantile(estimate, 0.975, na.rm = TRUE),
    boot_sd = sd(estimate, na.rm = TRUE),
    usable_reps = sum(!is.na(estimate)),
    .groups = "drop"
  ) %>%
  left_join(baseline_summary, by = "model")

boot_summary <- boot_summary_raw %>%
  mutate(
    Estimate = sprintf("%.3f", estimate),
    `Asymptotic 95\\% CI` = sprintf("[%.3f, %.3f]", asym_lo, asym_hi),
    `Bootstrap 95\\% CI` = sprintf("[%.3f, %.3f]", boot_lo, boot_hi),
    `Bootstrap SD` = sprintf("%.3f", boot_sd),
    `Usable reps` = usable_reps
  ) %>%
  select(model, Estimate, `Asymptotic 95\\% CI`, `Bootstrap 95\\% CI`, `Bootstrap SD`, `Usable reps`)

boot_kbl <- kable(
  boot_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Bootstrap uncertainty for procurement extension coefficients",
  col.names = c("Model", "Estimate", "Asymptotic 95\\% CI", "Bootstrap 95\\% CI", "Bootstrap SD", "Usable reps"),
  align = c("l", "r", "l", "l", "r", "r"),
  escape = FALSE
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  )

boot_kbl
boot_kbl %>% save_kable("results/table/bootstrap_uncertainty_hw3.tex")

```

Table \@ref(tab:robustboot) compares the model-based and bootstrap uncertainty around the `distance` coefficient in the four procurement models. If the bootstrap intervals are materially wider than the asymptotic intervals, then the matched extension sample is not large enough for the default standard errors to tell the full uncertainty story.

```{r fig-robustboot, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.cap = "Asymptotic versus bootstrap confidence intervals for distance coefficients", fig.pos="H", fig.align="center", out.width = "100%"}

stopifnot(exists("baseline_summary"), exists("boot_summary_raw"))

# compare asymptotic and bootstrap intervals visually
boot_plot_df <- baseline_summary %>%
  select(model, estimate, asym_lo, asym_hi) %>%
  pivot_longer(cols = c(asym_lo, asym_hi), names_to = "bound", values_to = "value")

boot_plot_df <- bind_rows(
  baseline_summary %>%
    transmute(model, estimate, lo = asym_lo, hi = asym_hi, interval = "Asymptotic"),
  boot_summary_raw %>%
    transmute(model, estimate, lo = boot_lo, hi = boot_hi, interval = "Bootstrap")
) %>%
  mutate(
    model = factor(model, levels = c(
      "Donor contract share",
      "Donor contract count",
      "Log donation amount",
      "Total contracts"
    )),
    interval = factor(interval, levels = c("Asymptotic", "Bootstrap"))
  )

ggplot(boot_plot_df,
       aes(x = estimate, y = model, xmin = lo, xmax = hi, color = interval)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_pointrange(position = position_dodge(width = 0.55)) +
  labs(
    x = "Coefficient on President-Agency Distance",
    y = "Dependent variable",
    color = "Interval type"
  ) +
  scale_color_manual(values = c("Asymptotic" = "steelblue", "Bootstrap" = "salmon")) +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("results/figure/bootstrap_uncertainty_hw3.png", width = 7, height = 4.5, dpi = 300)

```

Figure \@ref(fig:fig-robustboot) visualizes the same comparison. The distance coefficient is most stable when both the asymptotic and bootstrap intervals remain on the same side of zero. Wider bootstrap intervals indicate that the small matched sample adds uncertainty not obvious from the default model output alone.

\newpage

## Robustness Check 2. Leave-One-Agency-Out Sensitivity

```{r robustloo, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Leave-one-agency-out sensitivity summary", out.width = "100%"}

stopifnot(exists("ext_panel"), exists("baseline_summary"))

# re-estimate the four models dropping one matched agency at a time
omit_ids <- sort(unique(ext_panel$agency.id))

loo_results <- map_dfr(omit_ids, function(drop_id) {
  d <- ext_panel %>% filter(agency.id != drop_id)
  fits <- fit_boot_models(d)
  tibble(
    omitted_agency = drop_id,
    model = c("Donor contract share", "Donor contract count", "Log donation amount", "Total contracts"),
    estimate = c(
      if (!is.null(fits$share)) coef(fits$share)[["distance"]] else NA_real_,
      if (!is.null(fits$count)) coef(fits$count)[["distance"]] else NA_real_,
      if (!is.null(fits$amount)) coef(fits$amount)[["distance"]] else NA_real_,
      if (!is.null(fits$contracts)) coef(fits$contracts)[["distance"]] else NA_real_
    )
  )
})

loo_summary <- loo_results %>%
  left_join(baseline_summary %>% select(model, full_est = estimate), by = "model") %>%
  group_by(model, full_est) %>%
  summarize(
    min_est = min(estimate, na.rm = TRUE),
    max_est = max(estimate, na.rm = TRUE),
    sd_est = sd(estimate, na.rm = TRUE),
    share_same_sign = mean(sign(estimate) == sign(full_est), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    `Full estimate` = sprintf("%.3f", full_est),
    `LOO range` = sprintf("[%.3f, %.3f]", min_est, max_est),
    `LOO SD` = sprintf("%.3f", sd_est),
    `Share same sign` = sprintf("%.2f", share_same_sign)
  ) %>%
  select(model, `Full estimate`, `LOO range`, `LOO SD`, `Share same sign`)

loo_kbl <- kable(
  loo_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Leave-one-agency-out sensitivity summary",
  align = c("l", "r", "l", "r", "r")
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position"),
    font_size = 9
  )

loo_kbl
loo_kbl %>% save_kable("results/table/leave_one_out_hw3.tex")

```

Table \@ref(tab:robustloo) summarizes how much the `distance` coefficient changes when one agency is omitted at a time. The most reassuring case is when the range is narrow and the sign remains the same across nearly all leave-one-out estimates. Wider ranges and sign reversals indicate that the result is being carried by a small number of agencies.

```{r fig-robustloo, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.cap = "Leave-one-agency-out dispersion in distance coefficients", fig.pos="H", fig.align="center", out.width = "100%"}

stopifnot(exists("loo_results"), exists("baseline_summary"))

loo_plot_df <- loo_results %>%
  left_join(baseline_summary %>% select(model, full_est = estimate), by = "model") %>%
  mutate(
    model = factor(model, levels = c(
      "Donor contract share",
      "Donor contract count",
      "Log donation amount",
      "Total contracts"
    ))
  )

ggplot(loo_plot_df, aes(x = estimate, y = model)) +
  geom_vline(aes(xintercept = full_est), data = distinct(loo_plot_df, model, full_est),
             linetype = "dashed", color = "red") +
  geom_jitter(height = 0.12, width = 0, alpha = 0.65, color = "grey30") +
  labs(
    x = "Leave-one-agency-out coefficient on distance",
    y = "Dependent variable"
  ) +
  theme_bw()

ggsave("results/figure/leave_one_out_hw3.png", width = 7, height = 4.5, dpi = 300)

```

Figure \@ref(fig:fig-robustloo) shows the full distribution of leave-one-agency-out estimates. The dashed red line is the full-sample coefficient. Tight clusters around that line indicate stable results. Wide clouds or points crossing zero indicate that the small matched panel is sensitive to which agencies are included.

\newpage

## Robustness Check 3. Power-Style Simulation

```{r robustpower, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Power-style simulation for the matched procurement panel", out.width = "100%"}

stopifnot(exists("baseline_summary"), exists("ext_panel"))

# simulate retrospective power, Type S, and Type M errors for projected sample sizes
retro_sim <- function(beta_true, se_target, nsim = 5000, alpha = 0.05) {
  if (is.na(beta_true) || is.na(se_target) || se_target <= 0) {
    return(tibble(power = NA_real_, type_s = NA_real_, type_m = NA_real_))
  }
  z_crit <- qnorm(1 - alpha / 2)
  draws <- rnorm(nsim, mean = beta_true, sd = se_target)
  sig <- abs(draws / se_target) > z_crit
  power <- mean(sig)
  if (sum(sig) == 0 || abs(beta_true) < 1e-10) {
    type_s <- NA_real_
    type_m <- NA_real_
  } else {
    type_s <- mean(sign(draws[sig]) != sign(beta_true))
    type_m <- mean(abs(draws[sig] / beta_true))
  }
  tibble(power = power, type_s = type_s, type_m = type_m)
}

# evaluate projected power under two assumptions about the true effect size
n_current <- nrow(ext_panel)
n_targets <- c(48, 96, 192, 276)
assumption_scale <- c("Observed effect" = 1, "Half observed effect" = 0.5)

power_grid <- crossing(
  baseline_summary %>% select(model, estimate, std.error),
  assumption = names(assumption_scale),
  target_n = n_targets
) %>%
  mutate(
    scale = unname(assumption_scale[assumption]),
    beta_true = estimate * scale,
    se_target = std.error * sqrt(n_current / target_n)
  ) %>%
  rowwise() %>%
  mutate(stats = list(retro_sim(beta_true, se_target))) %>%
  unnest(stats) %>%
  ungroup()

# summarize current-sample power for the table
power_table <- power_grid %>%
  filter(target_n == n_current) %>%
  mutate(
    Power = sprintf("%.3f", power),
    `Type S` = ifelse(is.na(type_s), "NA", sprintf("%.3f", type_s)),
    `Type M` = ifelse(is.na(type_m), "NA", sprintf("%.3f", type_m))
  ) %>%
  select(model, assumption, Power, `Type S`, `Type M`)

power_kbl <- kable(
  power_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Power-style simulation for the matched procurement panel",
  col.names = c("Model", "Assumption", "Power", "Type S", "Type M"),
  align = c("l", "l", "r", "r", "r")
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position"),
    font_size = 9
  )

power_kbl
power_kbl %>% save_kable("results/table/power_simulation_hw3.tex")

```

Table \@ref(tab:robustpower) reports the power-style simulation at the current matched sample size. Power is shown under two assumptions: the observed HW2 effect size and a more conservative half-effect scenario. Type S is the probability of estimating the wrong sign conditional on statistical significance, and Type M is the exaggeration ratio conditional on significance. High Type M values indicate that significant results from a small sample are likely to overstate the true effect.

```{r fig-robustpower, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.cap = "Projected power as the matched sample size grows", fig.pos="H", fig.align="center", out.width = "100%"}

stopifnot(exists("power_grid"))

power_plot_df <- power_grid %>%
  mutate(
    model = factor(model, levels = c(
      "Donor contract share",
      "Donor contract count",
      "Log donation amount",
      "Total contracts"
    )),
    assumption = factor(assumption, levels = c("Observed effect", "Half observed effect"))
  )

ggplot(power_plot_df, aes(x = target_n, y = power, color = model)) +
  geom_hline(yintercept = 0.80, linetype = "dashed") +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_wrap(~ assumption) +
  scale_color_manual(values = c(
    "Donor contract share" = "#E69F00",
    "Donor contract count" = "#009E73",
    "Log donation amount" = "#56B4E9",
    "Total contracts" = "#CC79A7"
  )) +
  scale_x_continuous(breaks = n_targets) +
  labs(
    x = "Projected sample size",
    y = "Estimated power",
    color = "Dependent variable"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("results/figure/power_simulation_hw3.png", width = 8, height = 5, dpi = 300)

```

Figure \@ref(fig:fig-robustpower) traces how power changes as the matched panel grows. The dashed horizontal line marks the conventional 0.80 benchmark. If the lines remain well below 0.80 at `N = 48`, then the current extension is underpowered even before considering model uncertainty from agency matching.

\newpage

# 4. Results

## Table \@ref(tab:robustpanelsetup). Effective Sample Sizes Across ACS Replication and Extension Models

Table \@ref(tab:robustpanelsetup) shows the sample-size contrast motivating the HW3 robustness design. The original ACS replication uses 276 agency-administration observations, but the procurement extension relies on only 48 matched rows, while the ordered-logit follow-ups fall to 42 and 22 observations. This confirms that the main risk in the extension is not data abundance but limited information.

## Table \@ref(tab:robustboot). Bootstrap Uncertainty for Procurement Extension Coefficients

Table \@ref(tab:robustboot) shows that the bootstrap check does not affect the four procurement models in the same way. The donor-contract-share model remains centered at zero under both interval estimates, so this outcome provides little evidence of any distance effect. The donor-contract-count model stays positive, but both the asymptotic and bootstrap intervals include zero, which suggests directional stability without much precision. The clearest change appears in the log-donation model: the asymptotic interval excludes zero, but the bootstrap interval widens enough to include zero. That shift indicates that the negative HW2 estimate is substantively interesting but less certain once small-sample uncertainty is taken seriously. The total-contracts model stays negative, yet both intervals still include zero.

Substantively, the matched sample gives the clearest support for a negative relationship between ideological distance and the total amount donated within agency-administration cells. By contrast, the extension provides much weaker evidence that ideological distance changes donor-linked contract share, donor-linked contract count, or total contracting volume with comparable precision.

## Figure \@ref(fig:fig-robustboot). Asymptotic versus Bootstrap Confidence Intervals

Figure \@ref(fig:fig-robustboot) makes the same pattern easier to see. The bootstrap intervals for donor-contract share and donor-contract count are close to the asymptotic ones, but the log-donation and total-contract intervals widen noticeably. The log-donation result is still centered on a large negative coefficient, yet the bootstrap figure shows why that estimate should be interpreted as suggestive rather than definitive in a sample of 48 matched observations.

## Table \@ref(tab:robustloo). Leave-One-Agency-Out Sensitivity Summary

Table \@ref(tab:robustloo) shows that the leave-one-agency-out story is more reassuring than the bootstrap story for three of the four models. Donor-contract count stays positive in every omission, with a narrow range from 0.087 to 0.207. Total contracts also remains negative in every omission, and its range is fairly tight. The log-donation model keeps a negative sign throughout as well, although its range is wider, which means the coefficient is not driven by a single agency but its magnitude is sensitive to which agencies are in the sample. The donor-contract-share model is the weakest case. Its full-sample coefficient is essentially zero, and the sign changes in about 28 percent of the leave-one-out estimates.

## Figure \@ref(fig:fig-robustloo). Leave-One-Agency-Out Dispersion in Distance Coefficients

Figure \@ref(fig:fig-robustloo) reinforces that reading. The donor-contract-count and total-contract clouds are tightly clustered around the full-sample coefficient. The log-donation estimates are more dispersed and include a few visibly influential omissions, but they do not cross zero. By contrast, the donor-contract-share estimates sit almost directly on zero and move across it repeatedly, which is consistent with the absence of a stable signal in that model.

## Table \@ref(tab:robustpower). Power-Style Simulation for the Matched Procurement Panel

Table \@ref(tab:robustpower) shows that the current matched panel is not equally informative across outcomes. Under the observed-effect scenario, only the log-donation model clears the conventional 0.80 power threshold at `N = 48`. Total contracts reaches only moderate power (0.497), and donor-contract count remains low (0.245). Under the half-effect scenario, power drops sharply for every model. The donor-contract-share model is especially uninformative: because the observed coefficient is essentially zero, projected power stays at the nominal false-positive rate and the Type S and Type M summaries become mechanically unhelpful. The more informative comparison is between log donation, which remains relatively strong, and the count-based outcomes, which appear underpowered at the current sample size.

This matters substantively because the extension speaks most clearly to donation volume and much less clearly to downstream contracting counts. Mixed results in the count models should therefore be interpreted as limited evidence rather than as a sharp contradiction of the ideological-distance argument in either @acsmapping2025 or @fazekasagency2023.

## Figure \@ref(fig:fig-robustpower). Projected Power as the Matched Sample Size Grows

Figure \@ref(fig:fig-robustpower) shows how uneven that power problem is. With the observed HW2 effect sizes, log donation is already above 0.80 at `N = 48`, total contracts reaches 0.80 only after the sample doubles to about 96, and donor-contract count does not cross 0.80 until the sample is much closer to the original ACS size. Under the half-effect scenario, even total contracts remains below 0.80 at `N = 276`, and donor-contract count stays far below that benchmark throughout. The figure therefore supports a cautious interpretation of the count-based procurement results: mixed or null findings there may reflect low power rather than strong evidence of no relationship.

\newpage

# 5. Discussion

## Implications

The HW3 robustness checks shift the interpretation of the ACS-Fazekas extension from simple sign reading to a more careful assessment of what the sample can actually support. The original ACS replication remains strong because it is estimated on a much larger agency-administration panel. The procurement extension is substantively interesting, but it is built on a matched sample that is small enough for wide uncertainty, case sensitivity, and low power to become central inferential concerns.

The bootstrap and leave-one-agency-out checks answer complementary questions. The bootstrap asks how uncertain the coefficient is under repeated sampling from a small matched panel. The leave-one-agency-out analysis asks whether the result is being carried by a handful of agencies. In this application, the two checks point to a mixed conclusion rather than a simple rejection of the extension. The donor-contract-count and total-contract models are directionally stable when agencies are omitted one at a time, but the power analysis shows that both are still underpowered at the current sample size. The log-donation model is the strongest case substantively, yet the bootstrap indicates that its interval is wider than the model-based standard errors imply. The donor-contract-share model does not display a stable signal under any of the three checks.

The power-style simulation adds the final piece. It shows whether the observed sample size is large enough to detect effects of the size reported in HW2. The answer depends on the outcome. The log-donation model is reasonably well powered if the observed coefficient is close to the true one. The count-based outcomes are not. That means null or mixed findings in those models are weak evidence against the theory. The better conclusion is that the extension contains one relatively informative outcome, one near-null outcome, and two outcomes for which low power remains a serious constraint.

In substantive terms, the clearest pattern in the matched panel is that agencies farther from the president are associated with lower total donation volume. The matched sample is too small, however, to say with the same confidence whether ideological distance also changes donor-contract incidence or contract counts.

## Method-Theory Link

The theory in @acsmapping2025 is about ideological alignment between presidents and agencies. My HW2 extension translated that logic into procurement outcomes by asking whether the same ideological distance measure predicts donation and contracting behavior. HW3 does not introduce a new theory. Instead, it asks whether the extension is estimated precisely enough to support the substantive claims made from it. 

This is the relevant method-theory link for the homework: when the theory is tested on a much smaller matched sample than the original article used, the main threat is not only omitted variables or model choice. It is also whether the data contain enough information to distinguish signal from noise. 
In @acsmapping2025, ideological distance is meant to capture how far agencies sit from presidential preferences. In @fazekasagency2023, political contributions are one channel through which firms try to shape procurement relationships. The bootstrap, leave-one-agency-out, and power-style simulation all address whether those two ideas can be connected credibly in a small matched panel. Together they make the extension more informative because they separate weak theory from weak information.

## Limitations and Future Research

The main limitation of this HW3 design is that it treats the matched procurement panel as the object of inference without changing the underlying agency matching procedure. If the match itself is noisy, then the sample-size problem and the measurement problem are intertwined. A larger matched panel built from official agency identifiers would improve both coverage and statistical power.

A second limitation is that the power-style simulation is retrospective. It uses the observed HW2 estimates as anchors for plausible effect sizes. That is useful for diagnosing whether the current sample is informative, but it is not a substitute for a prospective design analysis based on a pre-specified substantive benchmark. 
A potential next step would be to define a minimum substantively important effect on donation or contracting outcomes and use that benchmark for a prospective power analysis.


\newpage

## References
