---
title: "PLSC 503: Homework 2"
subtitle: "Extension to Acs (2025): Mapping the Political Contours of the Regulatory State: Dynamic Estimates of Agency Ideal Points"
author: "Jody Park"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
colorlinks: true
urlcolor: blue
linkcolor: blue
bibliography: references.bib
csl: apanodoi.csl
link-citations: true
header-includes:
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{placeins}   
  - \usepackage{setspace}
  - \usepackage{dcolumn}
  - \usepackage{booktabs}
  - \usepackage{siunitx}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{enumerate}
  - \usepackage[shortlabels]{enumitem}
  - \usepackage[hang,flushmargin]{footmisc}	
  - \usepackage{url}
  - \usepackage{longtable}
  - \usepackage{graphicx}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{hyperref}
  - \hypersetup{
      colorlinks=true,
      linkcolor=blue,
      filecolor=magenta,      
      urlcolor=blue,
      citecolor=black,
    }
  - \usepackage[most]{tcolorbox}
  - \newtcolorbox{qbox}{enhanced,breakable,boxrule=0.5pt,arc=2mm,
    colback=gray!10,colframe=black!50,left=6pt,right=6pt,top=6pt,bottom=6pt}

---

```{r setup, include=FALSE}
options(knitr.table.format = "latex")
# clear all objects & free up memory if needed
rm(list = ls())

knitr::opts_chunk$set(
                      fig.width = 9, fig.height = 7, fig.retina = 3,
                      out.width = "100%",
                      cache = FALSE,
                      echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      hiline = TRUE
                      )
options(tinytex.verbose = TRUE)
knitr::opts_knit$set(allow_duplicate_labels = TRUE)

# set root directory to the project folder
# make sure whether the project folder exists in the folder
knitr::opts_knit$set(root.dir = normalizePath("."))

# check working directory
# getwd()

# load libraries
library(tidyverse)
library(ggpubr)
library(knitr)
library(kableExtra)
library(gridExtra)
library(tidyr)
library(haven)
library(readr)
library(caret)
library(modelr)
library(broom)
library(ggplot2)
library(plotly)
library(png)
library(stringr)
library(openxlsx)
library(Hmisc)
library(MASS) 
library(lfe) 
library(modelsummary) 
library(stargazer)
library(lubridate) 
library(scales) 
library(rdrobust) 
library(psych) 
library(nFactors)
library(here)
library(rstanarm)
library(loo)
library(xtable)

# resolve conflicts
library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::summarize)
conflicts_prefer(dplyr::summarise)
conflicts_prefer(dplyr::rename)
conflicts_prefer(dplyr::lag)
conflicts_prefer(psych::describe)
conflicts_prefer(dplyr::recode)

select <- dplyr::select
filter <- dplyr::filter
summarize <- dplyr::summarize
summarise <- dplyr::summarise
rename <- dplyr::rename
lag <- dplyr::lag
describe <- psych::describe
recode <- dplyr::recode


# Acs (2025) replication setup
options(scipen=100)
options(digits=2)

# setwd("acs-2025-apsr")

#### ------------ PREAMBLE SCRIPTS (OPTIONAL) -------------------
## rebuild CSV files used in main results
# install.packages("XML")
# source('code/preamble_oira.R', local = TRUE)
# source('code/preamble_ua.R', local = TRUE)
# source('code/preamble_merge.R', local = TRUE)

# Fazekas et al. (2023) replication setup
# load the authors' constants in the replication code
source("../fazekas-ferrali-wachs-2023-jpart/functions.R")
nRep <- 1e3
set.seed(503) 
pdf <- TRUE # set to TRUE to output manuscript figures to pdf
eps <- TRUE # set to TRUE to output manuscript figures to eps

# set base size for figures
theme_set(theme_minimal(base_size = 14))

```


# 1. Replication Setup

## Citations and Links

- **Original Paper**: ACS, A. (2025). Mapping the Political Contours of the Regulatory State: Dynamic Estimates of Agency Ideal Points. American Political Science Review, 1-18.

- **Replication Data**: [R replication codes and dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/C7S2SI) from @acsmapping2025 "Mapping the Political Contours of the Regulatory State: Dynamic Estimates of Agency Ideal Points (APSR)."

- **Paper for the Extension**: Fazekas, M., Ferrali, R., & Wachs, J. (2023). Agency Independence, Campaign Contributions, and Favoritism in US Federal Government Contracting. Journal of Public Administration Research and Theory, 33(2), 262-278.

- **Replication Data**: [R replication codes and dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/3U07EE) from @fazekasagency2023 "Agency Independence, Campaign Contributions, and Favoritism in US Federal Government Contracting (JPART)." 

- **Repository Links**:  

  - [Dropbox folder](https://www.dropbox.com/scl/fo/6b5w3zznokywvmlox3dt2/ANIInPxRVHuL4ys3Jqr0JX0?rlkey=3a2d8uvvkcfmbxoldhr91jobu&st=ucxm71yq&dl=0); current code and data for this paper are in the *acs-2025-apsr* subfolder.

  - [GitHub repository for this project](https://github.com/jodypark/plsc503_replication_extension_fazekas2023/tree/main/acs-2025-apsr)

\bigskip

## Article Introduction

@acsmapping2025 investigates the political ideology of US federal regulatory agencies by estimating dynamic ideal points for 69 administrative units across four presidential administrations (Clinton to Trump). The study uses data from the Unified Agenda on planned regulations and the president's discretionary review of those regulations from the Office of Information and Regulatory Affairs (OIRA) docket. 
By analyzing the ideological positions of regulatory agencies over time, the paper provides insights into how agency ideology shapes regulatory behavior and how it interacts with political dynamics. The core contribution of the paper is to offer a comprehensive mapping of the political contours of the regulatory state, shedding light on the ideological underpinnings of regulatory activity and its implications for policy outcomes.

I extend the analysis by examining how agency ideology moderates the effects of political party contributions on favoritism in federal contracting, using the dynamic ideal point estimates from @acsmapping2025. Merging this agency ideology data with @fazekasagency2023 allows a direct test of whether ideological alignment between the president and an agency amplifies the returns to political contributions in procurement.


\bigskip

## Data, Variables and Models

### Data

In @acsmapping2025, the author uses publicly available datasets:

1. **Unified Agenda data**: Includes information on the regulations that federal agencies plan to implement, including their content, timing, and regulatory impact (a semi-annual catalog, indicating pending or completed).

2. **The Office of Information and Regulatory Affairs (OIRA) docket**: Lists the regulations reviewed. For each regulation, the docket includes information on the agency responsible, the date of review, and the outcome of the review (e.g., approved, disapproved, or returned for revision).

3. **Federal Register data**: Contains the official publication of federal regulations and related documents. [Federal Register API](https://www.federalregister.gov/developers/documentation/api/v1)


@acsmapping2025 uses the eight-digit regulation identifier number (the RIN) to merge the Unified Agenda data with the OIRA docket data.

- The analysis focuses on 69 administrative units (e.g., EPA, FDA, etc.) from 2001 to 2020, covering four presidential administrations (Clinton, Bush, Obama, Trump). 

  - The author excludes regulatory agencies that "did not issue a regulation in each administration or issued fewer than 40 regulations during the entire period (averaging less than 10 per administration)" [@acsmapping2025].
  
  - This exclusion criterion ensures that the analysis focuses on regulatory agencies with sufficient regulatory activity to estimate meaningful ideal points and examine their dynamics over time.
  

### Variables

1. **Dependent Variable**:

- (1) `Review` (`Y_i`) (for the dynamic measurement model): Binary indicator for whether a planned rule in the Unified Agenda was reviewed by OIRA (`1` = reviewed at least once in the OIRA docket; `0` = not reviewed).

- (2) `Regulatory Activity_{at}` (for regulatory productivity model): Agency-level regulatory output in administration `t`, measured using Federal Register final-rule page length (average per year within administration, with alternative coding rules reported in Table 2 and Supplementary Appendix).


2. **Independent Variable**: 

- `President-Agency ideological distance`: `|X_p - X_{at}|`, where `X_p` is the president's fixed ideal point (DW-NOMINATE anchors) and `X_{at}` is the agency ideal point in administration `t`.

3. **Model Components**:

- `Rule significance` (`S_i`): Binary indicator from Unified Agenda priority categories (`1` = economically significant/significant; `0` = otherwise), plus interaction with spatial distance in the measurement model.

- *Model structure*: `Agency intercepts` (`a_a`) and `dynamic structure`: Agency-specific baseline review propensity and a random-walk evolution of agency ideal points across administrations (Clinton, Bush, Obama, Trump), estimated in Stan (Supplementary Appendix B).


\newpage


### Models

For this homework, I will focus on *Model 2* (Regulatory Productivity) and *H3*, which use a negative binomial model to examine the relationship between President-Agency ideological distance and regulatory output. 

- **Model 1 (Dynamic Measurement Model / IRT-Probit)**: The core model estimates agency ideal points from OIRA review decisions using a Bayesian probit IRT structure with president-agency distance, significance terms, agency intercepts, and administration-specific agency ideal points that evolve via a random walk. The Stan specification and priors are provided in Supplementary Appendix B.

- **Model 2 (Validation Model: Regulatory Productivity)**: The paper then estimates a *negative binomial model* (Equation 3 in the original paper) where regulatory activity is regressed on President-Agency Distance with agency fixed effects and multiple outcome constructions (Table 2), testing whether ideologically closer agencies are more productive.

$$
\text{Regulatory Activity}_{at} \sim \text{NegBin}(\mu_{at}, \theta)
$$
$$
\ln(\mu_{at}) = \pi_a + \beta \cdot \text{PresidentAgencyDistance}_{at}
\tag{3}
$$


**Tested Hypotheses in the Original Paper**:

- **H1 (Review-distance expectation)**: OIRA review probability increases as ideological distance between the president and an agency increases.

- **H2 (Dynamic ideology expectation)**: Agency ideology is not static; agency ideal points shift across presidential administrations.

- **H3 (Productivity-distance expectation)**: Regulatory productivity is lower when President-Agency ideological distance is larger (negative association in Equation 3 / Table 2).


\newpage


# 2. Replication Accuracy

## Figure 4. Conservative Trends in Agency Ideal Points

```{r f4repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, floating = FALSE, fig.pos="H", fig.align="center", fig.cap = "Conservative Trends in Agency Ideal Points", out.width = "100%"}

# use the author's code for figure 4: ---

####  FIGURE 4 (conservative trends)
x <- read.csv(file = 'data/dynamic_estimates.csv')
x <- subset(x, x$agency != 'BENCHMARK')

## average ideal points by admin
##  - statistics used in section 'Temporal Dynamics of Agency Ideal Points'
admin_means <- tapply(x$est, x$admin, mean)
## weighted.avg <- function(var, se) sum(var * (1/se^2)) / sum(1/se^2)
## tapply(x$est, x$admin, weighted.avg, x$se)
## (.04 - (-.11)) / abs(-.11)

z <- split(x, x$agency.id)
m <- lapply(z, function(k) {
    k$est.max  <- max(k$est)
    k$est.min  <- min(k$est)
    k$admin.max <- k$admin[k$est == k$est.max]
    k$admin.min <- k$admin[k$est == k$est.min]
    k$dif <- sqrt(abs(k$est.max - k$est.min))
    return(k)
})
r <- do.call('rbind', m)
rownames(r) <- 1:nrow(r)
r <- r[!duplicated(r$agency.id), c('agency.id', 'est.max', 'est.min', 'admin.max', 'admin.min', 'dif')]
r <- r[order(r$dif, decreasing = TRUE),]
rownames(r) <- 1:nrow(r)

## collect biggest shifters (top 10)
a <- r$agency.id[1:10]
y <- x[x$agency.id %in% a,]
## identify 7 conservative shifts from top 10
a <- c('AID', 'TREAS-FINCEN', 'DOD-DARC', 'OPM', 'DOD-DODOASHA', 'VA', 'EPA-AR')

y$administration <- factor(y$time.period,
                              levels = c(1, 2, 3, 4),
                              labels = c("Clinton", "Bush", "Obama", "Trump"))
y <- y[y$agency.id %in% a,]

## recode names
y$group[y$agency.id == 'OPM' | y$agency.id == 'EPA-AR' |
        y$agency.id == 'AID'] <- 'Shift During \n Trump Administration'
y$group[y$agency.id == 'VA' | y$agency.id == 'DOD-DODOASHA'] <- 'Gradual Trend \n Liberal Agencies'
y$group[y$agency.id == 'TREAS-FINCEN'
        | y$agency.id == 'DOD-DARC'] <- 'Gradual Trend \n Conserv. Agencies'
y$group <- factor(y$group,
                  levels = c('Gradual Trend \n Liberal Agencies',
                             'Shift During \n Trump Administration',
                             'Gradual Trend \n Conserv. Agencies'))

y$agency.name <- y$agency.id
y$agency.name[y$agency.id == 'EPA-AR'] <- 'EPA (Office of Air)'
y$agency.name[y$agency.id == 'DOD-DODOASHA'] <- 'Health Affairs (DOD-OASHA)'
y$agency.name[y$agency.id == 'OPM'] <- 'Personnel Management (OPM)'
y$agency.name[y$agency.id == 'VA'] <- 'Veterans Affairs (VA)'
y$agency.name[y$agency.id == 'AID'] <- 'International Development (AID)'
y$agency.name[y$agency.id == 'DOD-DARC'] <- 'Defense Acquisition (DOD-DARC)'
y$agency.name[y$agency.id == 'TREAS-FINCEN'] <- 'Financial Crimes (TREAS-FINCEN)'

f4rep <- ggplot(y, aes(x = administration, y = est, linetype = agency.name, shape = agency.name, color = agency.name, group = agency.name)) +
  geom_line() +
   theme(plot.margin = unit(c(0.1, 0.5, 0.5, 0.5), "cm")) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "red") +
    facet_wrap(~ group) +
  scale_linetype_manual(values = 1:7) +
  scale_shape_manual(values = 1:7) +
    scale_color_manual(values = c("salmon", "salmon", "darkgreen", "salmon",
                                  "steelblue", "darkgreen", "steelblue")) +
  labs(title = "",
       x = "",
       y = "",
       linetype = "Agency",
       shape = "Agency",
       color = "Agency") +
  theme_minimal(base_size = 14) +
  theme(panel.grid.major = element_line(linewidth = 0.5, colour = "grey"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.text = element_text(size = 8),
	        axis.text.x = element_text(angle = 45, hjust = 1)) +
	    guides(linetype = guide_legend(nrow = 4, title = NULL),
	           shape = guide_legend(nrow = 4, title = NULL),
	           color = guide_legend(nrow = 4, title = NULL))

print(f4rep)

# optional file export (keep the author's code)
dir.create("results", showWarnings = FALSE, recursive = TRUE)
dir.create("results/figure", showWarnings = FALSE, recursive = TRUE)
ggsave("results/figure4.pdf", plot = f4rep, width = 10, height = 7, units = "in")
ggsave("results/figure/f4repacs.png", plot = f4rep, width = 10, height = 7, dpi = 300)

```

Figure \@ref(fig:f4repacs) replicates the original pattern that several agencies shift in a conservative direction over administrations, with the largest movement concentrated in a small subset of agencies. This visual replication is consistent with @acsmapping2025's Figure 4.

\FloatBarrier

\newpage


## Table 1. Agencies with Largest Shifts in the Conservative Direction

```{r t1repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Agencies with Largest Shifts in the Conservative Direction", out.width = "100%"}

# use the author's code for table 1: ---

####  TABLE 1 (ideal point changes - follows from Figure 4)

# original code:
# ## use 'y' object from Figure 4
# 
# # Compute min and max by agency
# min.vals <- tapply(y$est, y$agency.id, min)
# max.vals <- tapply(y$est, y$agency.id, max)
# 
# # Combine into a data frame
# tab <- data.frame(
#   Agency = names(min.vals),
#   Min = as.numeric(min.vals),
#   Max = as.numeric(max.vals),
#   row.names = NULL
# )
# 
# # Create LaTeX table (NOTE: table is formatted differently in the paper)
# print(
#   xtable(tab,
#          align = c("l", "l", "r", "r")),
#   include.rownames = FALSE,
#   file = 'results/table1.tex')
# 


# use the original workflow object from Figure 4
stopifnot(exists("y"))

# Compute min and max by agency (original code logic)
min.vals <- tapply(y$est, y$agency.id, min)
max.vals <- tapply(y$est, y$agency.id, max)
tab <- data.frame(
  agency.id = names(min.vals),
  Min = as.numeric(min.vals),
  Max = as.numeric(max.vals),
  row.names = NULL
)

# paper order + labels/drivers
table1_labels <- tribble(
  ~agency.id,       ~Agency,                                  ~`Plausible Driver of Change`,                                ~ord,
  "OPM",            "Personnel Management (OPM)",             "\\makecell[l]{\\textit{Politicization:}\\\\Trump administration}",           1,
  "EPA-AR",         "EPA's Office of Air (EPA-AIR)",         "Trump administration",                                        2,
  "AID",            "Agency for Intl. Development (USAID)",  "Trump administration",                                        3,
  "VA",             "Veterans Affairs (VA)",                 "\\makecell[l]{\\textit{New policy conflict:}\\\\VA MISSION Act of 2018, wars}", 5,
  "DOD-DARC",       "Defense Acquisition (DOD-DARC)",        "Iraq/Afghan wars",                                            6,
  "TREAS-FINCEN",   "Financial Crimes (TREAS-FINCEN)",       "Patriot Act of 2001, civil liberties",                        7,
  "DOD-DODOASHA",   "Health Affairs, Def. (DOD-OASHA)",      "Iraq/Afghan wars",                                            8
)

table1_body <- table1_labels %>%
  left_join(tab, by = "agency.id") %>%
  mutate(
    Min = sprintf("%.2f", Min),
    Max = sprintf("%.2f", Max)
  ) %>%
  dplyr::select(ord, Agency, Min, Max, `Plausible Driver of Change`)

# spacer row between the two groups
table1_body <- bind_rows(
  table1_body,
  tibble(
    ord = 4,
    Agency = " ",
    Min = " ",
    Max = " ",
    `Plausible Driver of Change` = " "
  )
) %>%
  arrange(ord) %>%
  dplyr::select(-ord)

t1rep <- knitr::kable(
  table1_body,
  format = "latex",
  booktabs = TRUE,
  caption = "Agencies with Largest Shifts in the Conservative Direction",
  col.names = c("Agency", "Min", "Max", "Plausible Driver of Change"),
  align = c("l", "r", "r", "l"),
  escape = FALSE,
  linesep = ""
) %>%
  add_header_above(c(" " = 1, "Ideal Points" = 2, " " = 1), escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position")
  )

t1rep
t1rep %>% save_kable("results/table/t1repacs.tex")

```

Table \@ref(tab:t1repacs) reproduces the agency-level minima and maxima from the Figure 4 subset and highlights the same conservative movers. The ordering and substantive interpretation follow the original table.



\newpage


## Table 2. Ideological Distance and Regulatory Productivity (Negative Binomial)

```{r t2repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", out.width = "100%"}

# use the author's code for table 2: ---

####  TABLE 2 (productivity in number of rules)

# library(stargazer)
# library(lubridate)
# library(MASS)

v <- read.csv(file = 'data/estimation_data.csv')

v$agency.id[v$agency.abb == v$bureau.abb] <- v$agency.abb[v$agency.abb == v$bureau.abb]
v$agency.code <- substr(v$rin, 1, 4)
v <- v[!duplicated(v$agency.code), c('agency.id', 'agency.code')]
rownames(v) <- 1:nrow(v)
v$agency.code <- sprintf("%04s", v$agency.code)

y <- read.csv(file = 'data/dynamic_estimates.csv')

v$agency.id[!v$agency.id %in% y$agency.id] <- gsub('(.*)-.*', '\\1', v$agency.id[!v$agency.id %in% y$agency.id])

y <- merge(y, v, by = 'agency.id')
n_agency_admin <- nrow(y) # 276

## import OIRA data to get RINs of reviewed rules
load('data/oira_data/oira_data.RData')
g <- oira.data
colnames(g) <- c('agency.code', 'rin', 'rule', 'stage', 'econ.sig', 'date.received',
                 'legal.dline', 'date.completed', 'decision', 'date.published')
## recode variables
g$returned <- ifelse(g$decision == 'Returned for Reconsideration' |
                     g$decision == 'Withdrawn', 1, 0)
g$accepted <- ifelse(g$decision == 'Consistent without Change', 1, 0)
g$review.duration <- as.numeric(as.Date(g$date.completed) - as.Date(g$date.received))

r <- subset(g, date.received > '1995-01-01' & date.received < '2021-01-19')
r <- subset(r, stage == 'Proposed Rule')
## drop odd cases
d <- c('Exempt from Executive Order', 'Improperly Submitted', 'Emergency',
       'Statutory or Judicial Deadline')
r <- subset(r, !decision %in% d)
r$accepted <- ifelse(r$decision == 'Consistent without Change', 1, 0)
r$total <- rep(1, nrow(r))
r$admin <- rep('Clinton', nrow(r))
r$admin[as.Date(r$date.received) > '2001-01-19'] <- 'Bush'
r$admin[as.Date(r$date.received) > '2009-01-19'] <- 'Obama'
r$admin[as.Date(r$date.received) > '2016-01-19'] <- 'Trump'

r1 <- aggregate(cbind(returned, accepted, total) ~ agency.code + admin, data = r, sum, na.rm = TRUE)
r2 <- aggregate(review.duration ~ agency.code + admin, data = r, mean, na.rm = TRUE)
r1 <- merge(r1, r2, by = c('agency.code', 'admin'), all.x = TRUE)

v.expand <- expand.grid(agency.id = v$agency.id,
                    admin = c('Clinton', 'Bush', 'Obama', 'Trump'))
v.expand <- merge(v, v.expand, by =  'agency.id')
r1 <- merge(v.expand, r1, by = c('agency.code', 'admin'), all.x = T)
r1[is.na(r1)] <- 0 ## agencies with no reviews get count of 0

r3 <- aggregate(review.duration ~ agency.code + admin,
                data = subset(r, econ.sig == 'Yes'), mean, na.rm = TRUE)
colnames(r3)[3] <- 'review.dur.econ'
r <- merge(r1, r3, by = c('agency.code', 'admin'), all.x = TRUE)

## federal register data on final rules
x <- read.csv(file = 'data/final_rules_from_api.csv')

## fix date and admin
x$date <- mdy(x$publication_date)
x$date <- format(x$date, "%Y-%m-%d")
x$admin <- rep('Clinton', nrow(x))
x$admin[as.Date(x$date) > '2001-01-19'] <- 'Bush'
x$admin[as.Date(x$date) > '2009-01-19'] <- 'Obama'
x$admin[as.Date(x$date) > '2017-01-19'] <- 'Trump'
x$year <- substr(x$date, 1, 4)
x <- subset(x, date < '2021-01-20')

sig_table_raw <- table(x$significant)
x$sig <- x$significant
x$sig[x$significant == 'true'] <- 1
x$sig[x$significant == 'false'] <- 0
sig_table_binary <- table(x$sig)

x$rin <- gsub('^(.*?);.*', '\\1', x$regulation_id_numbers)

x$agency.code <- substr(x$rin, 1, 4)
x <- subset(x, !is.na(x$agency.code))
## counter for the number of rules
x$counter <- rep(1, nrow(x))

## fix action
n_action_raw <- length(unique(x$action))
x$action.new <- x$action
x$action.new <- gsub('ACTION:', '', x$action.new)
x$action.new <- trimws(x$action.new)
x$action.new <- gsub('  +', ' ', x$action.new)
x$action.new <- gsub('\\.$', '', x$action.new)
x$action.new <- tolower(x$action.new)
n_action_clean <- length(unique(x$action.new))

## categories to isolate final rules
x$action.new[grep('rescission|removal|commentary|temporary|technical amend|notice of enforcement|bridge|catch|final special conditions|withdraw|quota|immediate final rule|notice of availability|closure|clarification|sunset review|retention limits|delay|information collection|deferral|petition|correct|clarification|conditional|confirmation|compliance date|compliance deadline|extension|extend|expiration|change of|approval of|request for|court order|adjustments|modification|stay|interim|direct|information collection|advisory opinion|^notice$|interpret|notification of|notice of.*date|small entity compliance guide|meeting|guid', x$action.new)] <- 'direct.final.etc'
n_action_filtered <- length(unique(x$action.new))

## ASSUMPTION: what counts as a final rule
## drop.direct.rules = 0
drop.direct.rules = 1
if(drop.direct.rules == 0) {
    x$final.rule <- rep(1, nrow(x))
} else {
    x$final.rule <- rep(0, nrow(x))
    x$final.rule[grep('final rule|final reg|final order|final policy', x$action.new)] <- 1
}
final_rule_table <- table(x$final.rule)

## search for deregulatory action
## keywords associated with deregulation
keywords <- c("remove", "eliminate", "repeal", "reduce", "lower", "decrease",
              "simplify", "streamline", "ease", "exempt", "exclusion", "waive",
              "deregulate", "flexibility", "burden reduction", ## "flexibility act",
              "relief", "less burdensome", "sunset", "expiration", "reform", # "modernize"
              "cost savings", "efficiency", "remove barriers",
              "rescind", "decreased oversight", "less supervision", "voluntary compliance")
# Combine the keywords into a single regex pattern
pattern <- paste(keywords, collapse = "|")
x$dereg <- rep(0, nrow(x))
x$dereg[grep(pattern, x$abstract, ignore.case = TRUE)] <- 1
x$counter.regulatory <- x$counter
x$counter.regulatory[x$dereg == 1] <- 0
x$final.regulatory <- x$final.rule
x$final.regulatory[x$dereg == 1] <- 0

x$n.pages <- x$page_length
x$n.pages.regulatory <- x$page_length
x$n.pages.regulatory[x$dereg == 1] <- 0
x$n.pages.final.rule <- x$n.pages
x$n.pages.final.rule[x$final.rule == 0] <- 0

x$n.pages.final.rule.regulatory <- x$n.pages
x$n.pages.final.rule.regulatory[x$final.rule == 0] <- 0
x$n.pages.final.rule.regulatory[x$dereg == 1] <- 0

## significant
x$n.pages.final.rule.sig <- x$n.pages
x$n.pages.final.rule.sig[x$final.rule == 0] <- 0
x$n.pages.final.rule.sig[x$sig != 1] <- 0

## significant & regulatory
x$n.pages.final.rule.sig.reg <- x$n.pages
x$n.pages.final.rule.sig.reg[x$final.rule == 0] <- 0
x$n.pages.final.rule.sig.reg[x$sig != 1] <- 0
x$n.pages.final.rule.sig.reg[x$dereg == 1] <- 0
## x$n.pages.final.rule.sig[x$dereg == 1] <- 0

g <- g[g$stage == 'Proposed Rule',] # focus on reivew of proposal
x$n.pages.no.review <- x$n.pages  ## this version does not drop when x$final.rule == 0
x$n.pages.no.review[x$rin %in% unique(g$rin)] <- 0 # drop pages from reviewed rules
x$n.pages.no.review.final <- x$n.pages.no.review
x$n.pages.no.review.final[x$final.rule == 0] <- 0 # drop non-final rules
x$n.pages.no.review.final.dereg <- x$n.pages.no.review.final
x$n.pages.no.review.final.dereg[x$dereg == 1] <- 0 # drop de-regulatory rules

robust = 0
if(robust == 1) {
    x <- x[x$year != 2001 &
           x$year != 2009 &
           x$year != 2017,]
}

z <- aggregate(cbind(counter, counter.regulatory, final.rule, n.pages,
                     n.pages.final.rule, n.pages.final.rule.regulatory,
                     n.pages.regulatory, final.regulatory,
                     n.pages.final.rule.sig, n.pages.final.rule.sig.reg,
                     n.pages.no.review, n.pages.no.review.final,
                     n.pages.no.review.final.dereg) ~
               agency.code + admin, data = x, sum)
w <- aggregate(dereg ~ agency.code + admin, data = x, mean)
z <- merge(z, w, by = c('agency.code', 'admin'))

## y$agency.code %in% z$agency.code # all present

y <- merge(y, z, by = c('agency.code', 'admin'), all.x = T)
y[is.na(y)] <- 0

y$pres.ideal <- rep(.693, nrow(y))
y$pres.ideal[y$admin == 'Clinton'] <- -.438
y$pres.ideal[y$admin == 'Obama'] <- -.358
y$pres.ideal[y$admin == 'Trump'] <- .403
y$distance <- abs(y$pres.ideal - y$est)

year.adjust <- function(var, admin) {
    var[admin == 'Clinton'] <- var[admin == 'Clinton'] / 6 # (1995 - 2000)
    var[admin == 'Obama'] <- var[admin == 'Obama'] / 8
    var[admin == 'Bush'] <- var[admin == 'Bush'] / 8
    var[admin == 'Trump'] <- var[admin == 'Trump'] / 4
    return(var)
}

if(robust == 1) {
    year.adjust <- function(var, admin) {
        var[admin == 'Clinton'] <- var[admin == 'Clinton'] / 6 # (1995 - 2000)
        var[admin == 'Obama'] <- var[admin == 'Obama'] / 7
        var[admin == 'Bush'] <- var[admin == 'Bush'] / 7
        var[admin == 'Trump'] <- var[admin == 'Trump'] / 3
        return(var)
    }
}

y$counter.adj <- year.adjust(y$counter, y$admin)
y$counter.regulatory.adj <- year.adjust(y$counter.regulatory, y$admin)
y$final.rule.adj <- year.adjust(y$final.rule, y$admin)
y$final.regulatory.adj <- year.adjust(y$final.regulatory, y$admin)
y$n.pages.adj <- year.adjust(y$n.pages, y$admin)
y$n.pages.regulatory.adj <- year.adjust(y$n.pages.regulatory, y$admin)

y$n.pages.final.rule.adj <- year.adjust(y$n.pages.final.rule, y$admin)
y$n.pages.final.rule.regulatory.adj <- year.adjust(y$n.pages.final.rule.regulatory, y$admin)

y$n.pages.final.rule.sig.adj <- year.adjust(y$n.pages.final.rule.sig, y$admin)
y$n.pages.final.rule.sig.reg.adj <- year.adjust(y$n.pages.final.rule.sig.reg, y$admin)


y$n.pages.no.review.adj <- year.adjust(y$n.pages.no.review, y$admin)
y$n.pages.no.review.final.adj <- year.adjust(y$n.pages.no.review.final, y$admin)
y$n.pages.no.review.final.dereg.adj <- year.adjust(y$n.pages.no.review.final.dereg, y$admin)


n_obs_y <- nrow(y)
a <- aggregate(est ~ agency.id, data = y, mean)
colnames(a)[2] <- 'est.avg'
y <- merge(y, a, by = 'agency.id')
y$liberal.agency <- rep(0, nrow(y))
y$liberal.agency[y$est.avg < 0] <- 1

# keep a dedicated ACS panel object for later extension chunks
y_acs <- y

agency.fe <- c('Agency Fixed Effects', rep('$\\checkmark$', 5))

n <- c('President-Agency Distance')

# glm.nb() models for the five outcome constructions in Table 2, with distance and agency fixed effects, weighted by the inverse variance of each agency ideal point (1/se^2)
m1 <- glm.nb(as.integer(n.pages.final.rule.adj) ~ distance + agency.id, data = y, weights = 1/se^2)

m2 <- glm.nb(as.integer(n.pages.final.rule.sig.adj) ~ distance + agency.id, data = y, weights = 1/se^2)
m3 <- glm.nb(as.integer(n.pages.final.rule.sig.reg.adj) ~ distance + agency.id, data = y, weights = 1/se^2)

m4 <- glm.nb(as.integer(n.pages.no.review.final.adj) ~ distance + agency.id, data = y, weights = 1/se^2)

m5 <- glm.nb(as.integer(n.pages.no.review.final.dereg.adj) ~ distance + agency.id, data = y,
             weights = 1/se^2)

## exp(coef(m1)['distance']) - 1
## exp(coef(m2)['distance']) - 1
## exp(coef(m3)['distance']) - 1

options(digits=3)
dir.create("results", showWarnings = FALSE, recursive = TRUE)

# original code
# stargazer(m1, m2, m3, m4, m5,
#           ## type = 'text',
#           out = 'results/table2.tex',
#           omit = 'Constant|agency',
#           label = 'tbl:reg',
#           digits = 3,
#           title = 'Ideological Distance and Regulatory Productivity',
#           dep.var.labels = c('Formal', 'formal and sig', 'form, sig and Reg.', 'formal', 'formal and reg'),
#           star.cutoffs = c(.1, .05, .01),
#           star.char = c('^\\cdot', '*', '**'),
#           covariate.labels = n,
#           model.names = F,
#           add.lines = list(agency.fe),
#           no.space = TRUE,
#           font.size = 'small',
#           omit.stat=c("f", 'ser', 'adj.rsq', 'aic', 'bic', 'theta'))

# format coefficient and standard error for the distance row
fmt_t2 <- function(model) {
  s <- coef(summary(model))
  est <- unname(s["distance", "Estimate"])
  se  <- unname(s["distance", "Std. Error"])
  p   <- unname(s["distance", "Pr(>|z|)"])
  list(
    est = ifelse(
      !is.na(p) && p < 0.05,
      sprintf("$%0.3f^{*}$", est),
      sprintf("$%0.3f$", est)
    ),
    se  = sprintf("(%0.3f)", se)
  )
}

t2_stats <- lapply(list(m1, m2, m3, m4, m5), fmt_t2)
t2_est <- vapply(t2_stats, `[[`, character(1), "est")
t2_se  <- vapply(t2_stats, `[[`, character(1), "se")
t2_n   <- vapply(list(m1, m2, m3, m4, m5), nobs, numeric(1))

# keep the format like table 2 in the paper
t2rep <- c(
  "\\begin{table}[H]",
  "\\centering",
  "\\caption{Ideological Distance and Regulatory Productivity (Negative Binomial)}",
  "\\label{tab:t2repacs}",
  "\\small",
  "\\begin{tabular}{lccccc}",
  "\\hline\\hline",
  " & \\multicolumn{5}{c}{Dependent variable: Regulatory activity (final rule page length)} \\\\",
  "\\cline{2-6}",
  " & (1) & (2) & (3) & (4) & (5) \\\\",
  paste0(
    " & Formal & \\makecell[c]{Formal\\\\significant} & ",
    "\\makecell[c]{Formal significant\\\\regulatory} & ",
    "\\makecell[c]{Formal\\\\(excl. OIRA)} & ",
    "\\makecell[c]{Formal regulatory\\\\(excl. OIRA)} \\\\"
  ),
  "\\hline",
  paste0("Pres.--Agency Distance & ", paste(t2_est, collapse = " & "), " \\\\"),
  paste0(" & ", paste(t2_se, collapse = " & "), " \\\\"),
  "\\hline",
  "Agency Fixed Effects & $\\checkmark$ & $\\checkmark$ & $\\checkmark$ & $\\checkmark$ & $\\checkmark$ \\\\",
  paste0("Observations & ", paste(format(t2_n, trim = TRUE, scientific = FALSE), collapse = " & "), " \\\\"),
  "\\hline",
  paste0(
    "\\multicolumn{6}{l}{\\parbox{0.95\\textwidth}{$^{*}p < 0.05$. ",
    "Standard errors in parentheses. Models 4-5 exclude rules reviewed by OIRA at the proposed rule stage. ",
    "Models 3 and 5 also exclude deregulatory rules. Weights equal to the inverse variance of each agency ideal point are included.}} \\\\"
  ),
  "\\hline\\hline",
  "\\end{tabular}",
  "\\end{table}"
)

dir.create("results/table", showWarnings = FALSE, recursive = TRUE)
writeLines(t2rep, "results/table/t2repacs.tex")
cat(t2rep, sep = "\n")

```

Table \@ref(tab:t2repacs) replicates the negative-binomial validation result: larger President-Agency ideological distance is associated with lower regulatory productivity across outcome definitions. The sign and substantive magnitude patterns are aligned with the original paper.


\newpage


## Table A1. Agency Names and Ideal Point Estimates

```{r ta1repacs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Agency Names and Ideal Point Estimates", out.width = "100%"}

# use the author's code for table a1: ---

####   - TABLE A1

x <- read.csv(file = 'data/estimation_data.csv')
x <- subset(x, x$agency.id != 'BENCHMARK')

# get dept names
a <- read.csv(file = 'data/ua_data/agencycodes.csv')
a$agency.id <- paste(a$agency.abb, a$bureau.abb, sep = '-')
a <- subset(a, select = c(agency.id, agency, bureau, agency.abb, bureau.abb) )
a <- a[!duplicated(a),]
x <- merge(x, a, by = c('agency.id'), all.x = T)

x$agency.id.new <- x$agency.id
x$agency.id.new[x$agency.abb == x$bureau.abb] <- x$agency.abb[x$agency.abb == x$bureau.abb]

## create agency name table (APPENDIX TABLE 1)
x$agency.parent.new <- x$agency
x$agency.parent.new[x$agency.abb == x$bureau.abb] <- ''


y <- subset(x, select = c(agency.id.new, admin, sig, review))

#library(dplyr)
#library(tidyr)
#library(xtable)

collapsed_data <- y %>%
  group_by(admin, agency.id.new) %>%
  summarise(
    sig_mean = mean(sig, na.rm = TRUE),
    review_mean = mean(review, na.rm = TRUE),
    counter = n()
  ) %>%
  ungroup()

# convert to wide format
collapsed_pivot <- collapsed_data %>%
  pivot_wider(
    names_from = admin,
    values_from = c(sig_mean, review_mean, counter),
    names_glue = "{admin}_{.value}"
  )

z <- as.data.frame(collapsed_pivot)

# order the columns
ordered_columns <- c(
  "agency.id.new",
  "clinton_sig_mean", "clinton_review_mean", "clinton_counter",
  "bush43_sig_mean", "bush43_review_mean", "bush43_counter",
  "obama_sig_mean", "obama_review_mean", "obama_counter",
  "trump_sig_mean", "trump_review_mean", "trump_counter"
)

# reorder the dataframe columns
z <- z[, ordered_columns]

rownames(z) <- NULL
print(xtable(z),
      include.rownames = F,
      file = 'results/tableA1.tex')


# table a1 replication
# make display names latex-safe and readable in the knitted table
z_display <- z
colnames(z_display) <- c(
  "agency id",
  "clinton sig mean", "clinton review mean", "clinton n",
  "bush43 sig mean", "bush43 review mean", "bush43 n",
  "obama sig mean", "obama review mean", "obama n",
  "trump sig mean", "trump review mean", "trump n"
)

ta1rep_obj <- knitr::kable(
  z_display,
  format = "latex",
  booktabs = TRUE,
  caption = "Agency Names and Ideal Point Estimates",
  align = c("l", rep("c", ncol(z_display) - 1)),
  escape = TRUE
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  )

ta1rep_obj
ta1rep_obj %>% save_kable("results/table/ta1repacs.tex")
cat("\n\\FloatBarrier\n")

```

Table \@ref(tab:ta1repacs) reproduces the appendix-style agency-by-administration summaries of significance rates, review rates, and counts. These summaries support the data pipeline used to estimate the main replication models.


\newpage


# 3. Methodological Extension

## Why this method is necessary

The ACS validation model uses regulatory activity (final-rule page length) as the dependent variable to measure agency productivity. 

For this homework, I shift the dependent variable to agency outcomes from @fazekasagency2023, specifically donation and contracting outcomes, while retaining the ACS ideological distance measure as the main explanatory variable. 
This directly tests whether president-agency ideological alignment is associated with *who receives contracts* and *how politically active those contractors are*, not only with how many pages of rules are produced as a measure of agency productivity.

## How the method improves the paper

This extension keeps the ACS measure of president-agency ideological distance but changes the outcome from rule production to procurement outcomes. In the main extension models, I also add ACS-side controls for rule volume, significant rules, and review rates (`n_rules`, `n_sig_rules`, `review_rate`) so the estimated relationship is less likely to reflect simple differences in regulatory workload.

### Extension 1: Logistic Regression on Donation Probability

@acsmapping2025 establishes that ideological distance between the president and an agency depresses regulatory output. A parallel question on the firm side is whether donations flow toward ideologically close agencies or toward agencies where distance creates a greater demand for political access. I test this directly by regressing the binary donation indicator `isDonation` on `distance` using logistic regression with agency fixed effects, following Chapter 13 of @gelman2025:

$$
\operatorname{Pr}(\texttt{isDonation}_{ift} = 1) = \operatorname{logit}^{-1}(\alpha + \beta \cdot \texttt{distance}_{at} + \pi_a)
$$

where $\pi_a$ is an agency fixed effect and $\texttt{distance}_{at}$ is the ACS absolute president-agency ideological distance. A positive $\hat{\beta}$ indicates that firms donate more when ideological distance is larger. In this extension, that pattern is more consistent with compensatory or access-seeking donations to less aligned agencies than with simple targeting of already aligned agencies. This model requires a binomial distribution and logit link; the divide-by-4 rule translates the logit coefficient into an approximate marginal probability effect. The binary nature of `isDonation` makes the negative binomial distribution from ACS Equation (3) inappropriate, so logistic regression is the correct GLM for this outcome.


### Extension 2: Poisson versus Negative Binomial Overdispersion Diagnostic

Before extending the main results, I validate the negative binomial specification by reestimating ACS Equation (3) with a Poisson model and comparing fit using posterior predictive checks [@gelman2025]. The Poisson distribution is the nested special case of the negative binomial as the reciprocal dispersion parameter $\phi \to \infty$. 

For count outcomes, the Poisson model imposes $\text{sd}(y_i) = \sqrt{\text{E}(y_i)}$, which is violated whenever the data show more variance than this constraint allows. I generate 200 simulated replications $y^{\text{rep}}$ from both fitted models using `simulate()` on frequentist Poisson and negative binomial models fit with `glm` and `MASS::glm.nb`, and compare the frequency of zero counts in replicated data to the observed distribution. 

A Poisson model, which systematically underestimates zeros with replicated zero-rates far below the observed proportion, confirms that overdispersion is real and that the negative binomial specification in @acsmapping2025 is necessary. This diagnostic is independent of the main extension and applies directly to the replication sample.


### Extension 3: Ordered Logistic Regression on Agency Independence Tier

@acsmapping2025 estimates ideal points for agencies that span the full range of organizational independence. A remaining question is whether ideological distance predicts where an agency falls in @fazekasagency2023's four-tier independence hierarchy. I test this using ordered logistic regression, treating `agency_cat2` (1 = most independent, 4 = least independent) as an ordered categorical outcome:

$$
\operatorname{Pr}(\texttt{agency\_cat}_{a} > k) = \operatorname{logit}^{-1}(\beta \cdot \texttt{distance}_{a} - c_k), \quad k = 1, 2, 3
$$

where $c_k$ are cutpoints estimated alongside $\beta$. 

@fazekasagency2023's four-tier agency independence hierarchy:

- (1) `indep. commission & IRC`: most independent agencies

- (2) `indep. agency`: independent but not commission-based

- (3) `caninet/exec. dep. and bureau`: agencies with some independence but also more direct presidential oversight

- (4) `caninet/exec. dep., not bureau`: least independent agencies

A negative $\hat{\beta}$ implies that agencies with greater ideological distance from the president tend to occupy more independent tiers, consistent with the institutional logic that independent agencies face less direct presidential oversight. Here, the levels go from 1 (most independent) to 4 (least independent). 

I fit this using `MASS::polr`, which estimates $\beta$ and the cutpoints $c_k$ by maximum likelihood. This model is appropriate because the four tiers are ordered (independence is a latent continuous construct bucketed into discrete categories) but the gaps between tiers are not necessarily equal, which rules out treating `agency_cat` as a continuous dependent variable in OLS.


## Explanatory Tables and Figures

```{r extvardefs, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "ACS distance + Fazekas outcomes", out.width = "100%"}

# define variables used in the extension section
ext_var_defs <- tibble(
  Variable = c(
    "distance",
    "donor_contract_share",
    "donor_contract_count",
    "donation_amount",
    "n_contracts",
    "n_rules",
    "n_sig_rules",
    "review_rate"
  ),
  Role = c(
    "Independent variable (ACS)",
    "Dependent variable (Fazekas)",
    "Dependent variable (Fazekas)",
    "Dependent variable (Fazekas)",
    "Dependent variable (Fazekas)",
    "Control / alternative productivity measure (ACS)",
    "Control / alternative productivity measure (ACS)",
    "Control / alternative productivity measure (ACS)"
  ),
  Definition = c(
    "Absolute president-agency ideological distance from ACS dynamic estimates",
    "Share of contracts in agency-admin cell awarded to donor firms",
    "Count of donor contracts in agency-admin cell",
    "Total donation amount linked to firms in agency-admin cell",
    "Total number of contracts in agency-admin cell",
    "Number of planned rules in ACS estimation data",
    "Number of significant planned rules in ACS estimation data",
    "Mean OIRA review indicator in ACS estimation data"
  )
)

# print a compact variable-definition table
extvardefs_kbl <- kable(
  ext_var_defs,
  format = "latex",
  booktabs = TRUE,
  caption = "ACS distance + Fazekas outcomes",
  align = c("l", "l", "l")
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  )
extvardefs_kbl
extvardefs_kbl %>% save_kable("results/table/extvardefs.tex")

```

Table \@ref(tab:extvardefs) defines the merged-variable design used below. The dependent variables are now donation and contracting outcomes from @fazekasagency2023, while ACS ideological distance remains the core predictor.

```{r extbuildpanel, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "Merge diagnostics for ACS-Fazekas extension panel", out.width = "100%"}

# standardize agency names before matching
clean_agency_text <- function(x) {
  x %>%
    str_to_lower() %>%
    str_replace_all("&", " and ") %>%
    str_replace_all("\\badmin\\.?\\b", "administration") %>%
    str_replace_all("[^a-z0-9]+", " ") %>%
    str_squish()
}

# build ACS distance by agency and administration
acs_dynamic_ext <- read_csv("data/dynamic_estimates.csv", show_col_types = FALSE) %>%
  filter(agency.id != "BENCHMARK") %>%
  mutate(
    pres_ideal = case_when(
      admin == "Clinton" ~ -0.438,
      admin == "Bush" ~ 0.693,
      admin == "Obama" ~ -0.358,
      admin == "Trump" ~ 0.403,
      TRUE ~ NA_real_
    ),
    distance = abs(pres_ideal - est)
  ) %>%
  select(agency.id, admin, distance, est, se) %>%
  distinct()

# aggregate ACS-side productivity controls
acs_prod_ext <- read_csv("data/estimation_data.csv", show_col_types = FALSE) %>%
  mutate(
    sig = as.numeric(sig),
    review = as.numeric(review)
  ) %>%
  group_by(agency.id, admin) %>%
  summarize(
    n_rules = n(),
    n_sig_rules = sum(sig == 1, na.rm = TRUE),
    review_rate = mean(review, na.rm = TRUE),
    .groups = "drop"
  )

# prepare ACS agency names for matching
acs_name_ext <- read_csv("data/rcl_ideal_points_merged.csv", show_col_types = FALSE) %>%
  transmute(
    agency.id,
    agency_name_acs = coalesce(cleaned_agency, agency.y, agency.x)
  ) %>%
  filter(!is.na(agency_name_acs)) %>%
  distinct() %>%
  mutate(agency_name_clean = clean_agency_text(agency_name_acs)) %>%
  arrange(agency.id) %>%
  distinct(agency_name_clean, .keep_all = TRUE)

# recode Fazekas contracts to agency-admin outcomes
faz_ext <- read_csv("../fazekas-ferrali-wachs-2023-jpart/contracts.csv", show_col_types = FALSE) %>%
  mutate(
    donation_total = pmax(current_R_donor, 0) + pmax(current_D_donor, 0),
    donation_dummy = as.integer(donation_total > 0),
    admin = case_when(
      year <= 2008 ~ "Bush",
      year >= 2009 & year <= 2016 ~ "Obama",
      year < 2001 ~ "Clinton",
      TRUE ~ "Trump"
    ),
    agency_name_clean = clean_agency_text(agency_name)
  ) %>%
  filter(admin %in% c("Bush", "Obama"))

# merge the ACS and Fazekas pieces into one panel
ext_panel <- faz_ext %>%
  group_by(agency_name, agency_name_clean, admin) %>%
  summarize(
    n_contracts = n(),
    donor_contract_share = mean(donation_dummy, na.rm = TRUE),
    donor_contract_count = sum(donation_dummy, na.rm = TRUE),
    donation_amount = sum(donation_total, na.rm = TRUE),
    mean_log_donation = mean(log1p(donation_total), na.rm = TRUE),
    unique_firms = n_distinct(pduns),
    unique_donor_firms = n_distinct(pduns[donation_dummy == 1]),
    .groups = "drop"
  ) %>%
  left_join(
    acs_name_ext %>% select(agency_name_clean, agency.id),
    by = "agency_name_clean"
  ) %>%
  left_join(acs_dynamic_ext, by = c("agency.id", "admin")) %>%
  left_join(acs_prod_ext, by = c("agency.id", "admin")) %>%
  filter(!is.na(distance))

# merge coverage
ext_merge_check <- tibble(
  Metric = c(
    "Agency-admin rows in Fazekas aggregate (Bush/Obama)",
    "Rows matched to ACS distance",
    "Distinct matched agencies",
    "Mean donor contract share",
    "Mean contracts per matched agency-admin row"
  ),
  Value = c(
    nrow(faz_ext %>% distinct(agency_name_clean, admin)),
    nrow(ext_panel),
    n_distinct(ext_panel$agency.id),
    mean(ext_panel$donor_contract_share, na.rm = TRUE),
    mean(ext_panel$n_contracts, na.rm = TRUE)
  )
) %>%
  mutate(
    Value = ifelse(grepl("Mean", Metric), sprintf("%.3f", Value), as.character(round(Value, 0)))
  )

# merge diagnostics
extbuild_kbl <- kable(
  ext_merge_check,
  format = "latex",
  booktabs = TRUE,
  caption = "Merge diagnostics for ACS-Fazekas extension panel",
  align = c("l", "r")
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position")
  )
extbuild_kbl
extbuild_kbl %>% save_kable("results/table/extbuildpanel.tex")

```


Table \@ref(tab:extbuildpanel) reports the merge coverage for the new agency-administration panel. The matched panel is the estimation sample used in the extension models below.

```{r extmodeltable, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results='asis', floating = FALSE, tab.pos="H", tab.align="center", tab.cap = "ACS distance predicting Fazekas donation and contract outcomes", out.width = "100%"}

stopifnot(exists("ext_panel"), nrow(ext_panel) > 0)

# clean model inputs and fill missing controls
ext_panel <- ext_panel %>%
  mutate(
    n_contracts = pmax(n_contracts, 1),
    n_rules = replace_na(n_rules, 0),
    n_sig_rules = replace_na(n_sig_rules, 0),
    review_rate = replace_na(review_rate, 0)
  )

# estimate the main extension models
ext_mods <- list(
  "Donor contract share" = lm(
    donor_contract_share ~ distance + log1p(n_rules) + log1p(n_sig_rules) + review_rate + factor(admin),
    data = ext_panel
  ),
  "Donor contract count" = glm.nb(
    donor_contract_count ~ distance + log1p(n_rules) + log1p(n_sig_rules) + factor(admin) + offset(log(n_contracts)),
    data = ext_panel
  ),
  "Log donation amount" = lm(
    log1p(donation_amount) ~ distance + log1p(n_rules) + log1p(n_sig_rules) + review_rate + factor(admin),
    data = ext_panel
  ),
  # treat total contracts as a separate count outcome, not a control for donor-linked outcomes
  "Total contracts" = glm.nb(
    n_contracts ~ distance + factor(admin),
    data = ext_panel
  )
)

# relabel coefficients for the table
ext_coef_map <- c(
  "distance" = "President-Agency Distance",
  "log1p(n_rules)" = "Log(1 + number of rules)",
  "log1p(n_sig_rules)" = "Log(1 + significant rules)",
  "review_rate" = "OIRA review rate",
  "factor(admin)Obama" = "Obama administration (ref: Bush)"
)

# the extension regression table
ext_table <- modelsummary(
  ext_mods,
  coef_map = ext_coef_map,
  statistic = "({std.error})",
  stars = c("*" = 0.05),
  fmt = 3,
  gof_map = tibble::tribble(
    ~raw,   ~clean,     ~fmt,
    "nobs", "Num.Obs.", 0
  ),
  gof_omit = "AIC|BIC|Log.Lik.|F|RMSE|R2|Adj|Pseudo|Std",
  output = "latex",
  title = "ACS distance and Fazekas donation/contract outcomes",
  escape = FALSE,
  notes = "Standard errors in parentheses. Unit of analysis is the agency-administration cell.",
  notes_append = TRUE
) %>%
  add_header_above(c(" " = 1, "Dependent variables from Fazekas (2023)" = length(ext_mods)), escape = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  )

ext_table
ext_table %>% save_kable("results/table/extmodeltable.tex")

```

Table \@ref(tab:extmodeltable) extends ACS Equation (3) by replacing regulatory activity outcomes with donation and contracting outcomes from @fazekasagency2023. Columns 1 and 3 use `lm` for continuous outcomes, while columns 2 and 4 use `glm.nb` for count outcomes. The sign on `distance` is not uniform across outcomes, which suggests that ideological distance does not map onto every procurement outcome in the same way.


```{r extcoefplot, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.cap = "Distance effect across alternative Fazekas outcomes", fig.pos="H", fig.align="center", out.width = "100%"}

stopifnot(exists("ext_mods"))

# collect the distance coefficient from each model
ext_coef_plot <- bind_rows(
  broom::tidy(ext_mods[[1]], conf.int = TRUE) %>% mutate(model = "Donor contract share"),
  broom::tidy(ext_mods[[2]], conf.int = TRUE) %>% mutate(model = "Donor contract count"),
  broom::tidy(ext_mods[[3]], conf.int = TRUE) %>% mutate(model = "Log donation amount"),
  broom::tidy(ext_mods[[4]], conf.int = TRUE) %>% mutate(model = "Total contracts")
) %>%
  filter(term == "distance") %>%
  mutate(model = factor(model, levels = c(
    "Donor contract share",
    "Donor contract count",
    "Log donation amount",
    "Total contracts"
  )))

# point estimates with CIs
ggplot(ext_coef_plot, aes(x = estimate, y = model, xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_point() +
  geom_linerange() +
  labs(
    x = "Coefficient on President-Agency Distance",
    y = "Dependent variable"
  )
ggsave("results/figure/extcoefplot.png", width = 7, height = 4, dpi = 300)

```

Figure \@ref(fig:extcoefplot) compares the `distance` coefficient across multiple dependent-variable definitions. This directly addresses sensitivity to outcome choice, which is one of the main robustness goals for this homework.

\newpage

## Extension 1: Logistic Regression on Donation Probability

```{r ext1logit, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results = 'asis', floating = FALSE, tab.pos = "H", tab.align = "center", tab.cap = "Logistic Regression of Donation Probability on Ideological Distance", out.width = "100%"}

stopifnot(exists("faz_ext"), exists("acs_name_ext"), exists("acs_dynamic_ext"))

# merge contract outcomes with ACS distance
logit_data <- faz_ext %>%
  left_join(
    acs_name_ext %>% select(agency_name_clean, agency.id),
    by = "agency_name_clean"
  ) %>%
  left_join(acs_dynamic_ext, by = c("agency.id", "admin")) %>%
  filter(!is.na(distance), !is.na(donation_dummy))

# estimate baseline and admin-adjusted logits
m_logit1 <- glm(donation_dummy ~ distance,
                data = logit_data, family = binomial(link = "logit"))
m_logit2 <- glm(donation_dummy ~ distance + factor(admin),
                data = logit_data, family = binomial(link = "logit"))

# the logit regression table
t7ext <- modelsummary(
  list("(1) No controls" = m_logit1, "(2) Admin FE" = m_logit2),
  coef_map = c(
    "distance"           = "President-Agency Distance",
    "factor(admin)Obama" = "Obama (ref: Bush)"
  ),
  statistic = "({std.error})",
  stars     = c("*" = 0.05),
  fmt       = 3,
  gof_map   = tibble::tribble(
    ~raw,     ~clean,           ~fmt,
    "nobs",   "Observations",   0,
    "logLik", "Log Likelihood", 3
  ),
  gof_omit = "AIC|BIC|F|RMSE|R2|Adj|Pseudo|Std|Theta",
  output   = "latex",
  title    = "Logistic Regression of Donation Probability on Ideological Distance",
  escape   = FALSE,
  notes    = "Standard errors in parentheses. Unit of analysis is the contract.",
  notes_append = TRUE
) %>%
  kable_styling(
    full_width    = FALSE,
    position      = "center",
    latex_options = c("HOLD_position"),
    font_size     = 10
  )

t7ext
t7ext %>% save_kable("results/table/ext1logit.tex")

```

Table \@ref(tab:ext1logit) reports logistic regression coefficients for `donation_dummy` on `distance`. The coefficient on `distance` is positive in both specifications, indicating that firms are more likely to donate when ideological distance between the president and the agency is larger. This suggests that firms may use donations not only to reward aligned agencies, but also to secure access when ideological distance makes agency behavior less predictable.

```{r ext1logitplot, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.cap = "Predicted Probability of Donation by President-Agency Ideological Distance", fig.pos = "H", fig.align = "center", out.width = "80%"}

stopifnot(exists("m_logit2"), exists("logit_data"))

# create a prediction grid over the observed distance range
dist_seq <- seq(
  min(logit_data$distance, na.rm = TRUE),
  max(logit_data$distance, na.rm = TRUE),
  length.out = 100
)

# compute predicted probabilities by administration
pred_df <- bind_rows(
  tibble(distance = dist_seq, admin = "Bush"),
  tibble(distance = dist_seq, admin = "Obama")
) %>%
  mutate(pred = predict(m_logit2, newdata = ., type = "response"))

# plot the fitted probability curves
ggplot(pred_df, aes(x = distance, y = pred,
                    linetype = admin, color = admin)) +
  geom_line(linewidth = 0.9) +
  labs(
    x        = "President-Agency Ideological Distance",
    y        = "Predicted Pr(Donation = 1)",
    linetype = "Administration",
    color    = "Administration"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
ggsave("results/figure/ext1logitplot.png", width = 7, height = 5, dpi = 300)

```

Figure \@ref(fig:ext1logitplot) plots the predicted probability of donation as a function of `distance` for Bush and Obama administrations. The upward slope indicates that firms are more likely to donate as ideological distance increases.

\newpage

## Extension 2: Overdispersion Diagnostic

```{r ext2ppc, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.cap = "Posterior Predictive Check: Proportion of Zeros in Poisson vs. Negative Binomial Replications", fig.pos = "H", fig.align = "center", out.width = "90%"}

stopifnot(exists("y_acs"))

# prepare the ACS count outcome used in the original validation model
y_ppc <- y_acs %>%
  filter(!is.na(n.pages.final.rule.adj), !is.na(distance),
         !is.na(se), se > 0) %>%
  mutate(outcome = as.integer(round(n.pages.final.rule.adj)))

# fit Poisson and negative binomial versions of the same model
m_pois_ppc <- glm(
  outcome ~ distance + agency.id,
  data    = y_ppc,
  family  = poisson(link = "log"),
  weights = 1 / y_ppc$se^2
)

m_nb_ppc <- MASS::glm.nb(
  outcome ~ distance + agency.id,
  data    = y_ppc,
  weights = 1 / y_ppc$se^2
)

# simulate replicated datasets for the zero-rate comparison
set.seed(503)
nsims     <- 200
pois_sims <- simulate(m_pois_ppc, nsim = nsims)
nb_sims   <- simulate(m_nb_ppc,   nsim = nsims)

obs_zero  <- mean(y_ppc$outcome == 0, na.rm = TRUE)
pois_zero <- vapply(pois_sims, function(s) mean(s == 0, na.rm = TRUE), numeric(1))
nb_zero   <- vapply(nb_sims,   function(s) mean(s == 0, na.rm = TRUE), numeric(1))

# combine simulated zero-rates for plotting
ppc_df <- bind_rows(
  tibble(zero_rate = pois_zero, Model = "Poisson"),
  tibble(zero_rate = nb_zero,   Model = "Negative Binomial")
)

# figure: observed vs simulated zero-rates
ggplot(ppc_df, aes(x = zero_rate, fill = Model)) +
  geom_histogram(alpha = 0.6, bins = 25, position = "identity") +
  geom_vline(xintercept = obs_zero, linetype = "dashed", linewidth = 0.9) +
  scale_fill_manual(values = c("Poisson" = "salmon",
                               "Negative Binomial" = "steelblue")) +
  labs(
    x       = "Proportion of zeros in replicated data",
    y       = "Count of simulations",
    fill    = "Model",
    caption = "Dashed line: observed proportion of zeros in ACS data"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
ggsave("results/figure/ext2ppc.png", width = 7, height = 5, dpi = 300)

```

Figure \@ref(fig:ext2ppc) compares the proportion of zeros in 200 simulated datasets from Poisson and negative binomial models to the observed zero-rate (dashed line). The Poisson model systematically underestimates the proportion of zeros, while the negative binomial model generates replications closer to the observed rate, validating the negative binomial specification in @acsmapping2025.

\newpage

## Extension 3: Ordered Logistic Regression on Agency Independence Tier

```{r ext3ordlogit, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results = 'asis', floating = FALSE, tab.pos = "H", tab.align = "center", tab.cap = "Ordered Logistic Regression of Agency Independence Tier on Ideological Distance", out.width = "100%"}

stopifnot(exists("ext_panel"), exists("clean_agency_text"))

# recover agency independence categories from the Fazekas data
faz_cat_lookup <- read_csv(
  "../fazekas-ferrali-wachs-2023-jpart/contracts.csv",
  show_col_types = FALSE
) %>%
  mutate(agency_name_clean = clean_agency_text(agency_name)) %>%
  group_by(agency_name_clean) %>%
  summarize(agency_cat2 = first(agency_cat2), .groups = "drop") %>%
  filter(!is.na(agency_cat2))

# order categories from most to least independent
indep_levels <- c(
  "indep. commission & IRC",
  "indep. agency",
  "cabinet/exec. dep. and bureau",
  "cabinet/exec. dep. (not bureau)"
)

# merge categories onto the extension panel
ext_panel_cat <- ext_panel %>%
  left_join(faz_cat_lookup, by = "agency_name_clean") %>%
  filter(!is.na(agency_cat2), !is.na(distance)) %>%
  mutate(agency_cat_ord = factor(agency_cat2,
                                  levels = indep_levels,
                                  ordered = TRUE))

# estimate the ordered logit model
m_ordlogit <- MASS::polr(
  agency_cat_ord ~ distance,
  data   = ext_panel_cat,
  Hess   = TRUE,
  method = "logistic"
)

# the ordered logit table
t8ext <- modelsummary(
  list("Agency Independence Tier" = m_ordlogit),
  coef_map  = c("distance" = "President-Agency Distance"),
  coef_omit = "\\|",
  statistic = "({std.error})",
  stars     = c("*" = 0.05),
  fmt       = 3,
  gof_map   = tibble::tribble(
    ~raw,     ~clean,           ~fmt,
    "nobs",   "Observations",   0,
    "logLik", "Log Likelihood", 3
  ),
  gof_omit = "AIC|BIC|F|RMSE|R2|Adj|Pseudo|Std|Theta",
  output   = "latex",
  title    = "Ordered Logistic Regression of Agency Independence Tier on Ideological Distance",
  escape   = FALSE,
  notes    = "Standard errors in parentheses. Outcome ordered from most independent to least independent. Cutpoints omitted."
) %>%
  kable_styling(
    full_width    = FALSE,
    position      = "center",
    latex_options = c("hold_position", "scale_down"),
    font_size     = 9
  )

t8ext
t8ext %>% save_kable("results/table/ext3ordlogit.tex")

```

Table \@ref(tab:ext3ordlogit) reports the ordered logistic coefficient for `distance` on the four-tier agency independence scale, from most independent (level 1: indep. commission and IRC) to least independent (level 4: cabinet department). A negative $\hat{\beta}$ implies that agencies with greater ideological distance from the president tend to fall into more independent (lower-numbered) tiers, consistent with the institutional logic that structurally independent agencies face less direct presidential alignment pressure. A null result would indicate that ideological distance and formal independence are orthogonal properties in this sample.

## Extension 3a: Agency-Level Alternative

Extension 3a re-estimates Extension 3 after collapsing the panel to one row per agency, replacing administration-specific `distance` with each agency's mean ideological distance across administrations. The motivation is statistical: agency independence tier is a fixed agency-level characteristic that does not vary across administrations, so treating each agency-administration row as an independent observation inflates the effective sample size in Extension 3. The agency-level specification corrects this by using the natural unit of the outcome variable. If the coefficient on mean distance is negative and close in magnitude to Extension 3, the cross-administration result is more credible as a stable cross-agency pattern rather than an artifact of repeated observations.

```{r ext3aagencylevel, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results = 'asis', floating = FALSE, tab.pos = "H", tab.align = "center", tab.cap = "Agency-Level Ordered Logistic Regression of Independence Tier on Mean Ideological Distance", out.width = "100%"}

stopifnot(exists("ext_panel_cat"))

# collapse to one row per agency
ext_panel_agency <- ext_panel_cat %>%
  group_by(agency.id, agency_name_clean, agency_cat2) %>%
  summarize(
    mean_distance = mean(distance, na.rm = TRUE),
    n_admin = n(),
    .groups = "drop"
  ) %>%
  mutate(
    agency_cat_ord = factor(
      agency_cat2,
      levels = indep_levels,
      ordered = TRUE
    )
  )

# estimate the agency-level ordered logit
m_ordlogit_agency <- MASS::polr(
  agency_cat_ord ~ mean_distance,
  data   = ext_panel_agency,
  Hess   = TRUE,
  method = "logistic"
)

# table 9
t9extwide <- modelsummary(
  list("Agency Independence Tier" = m_ordlogit_agency),
  coef_map  = c("mean_distance" = "Mean President-Agency Distance"),
  coef_omit = "\\|",
  statistic = "({std.error})",
  stars     = c("*" = 0.05),
  fmt       = 3,
  gof_map   = tibble::tribble(
    ~raw,     ~clean,         ~fmt,
    "nobs",   "Observations", 0
  ),
  gof_omit = "AIC|BIC|F|RMSE|R2|Adj|Pseudo|Std|Theta|Log.Lik",
  output   = "latex",
  title    = "Agency-Level Ordered Logistic Regression of Independence Tier on Mean Ideological Distance",
  escape   = FALSE,
  notes    = "Standard errors in parentheses. Outcome ordered from most independent to least independent. One row per agency."
) %>%
  kable_styling(
    full_width    = FALSE,
    position      = "center",
    latex_options = c("hold_position", "scale_down"),
    font_size     = 9
  )

t9extwide
t9extwide %>% save_kable("results/table/ext3aagencylevel.tex")

```
Table \@ref(tab:ext3aagencylevel) reports the same ordered-logit idea as Table \@ref(tab:ext3ordlogit), but with the unit of analysis collapsed to the agency level. This version is substantively closer to the institutional outcome because agency independence tier does not vary across administrations. A negative coefficient on mean distance implies that agencies further from the president tend to fall into more independent tiers.

\FloatBarrier

\newpage

# 4. Results

## Figure \@ref(fig:f4repacs). Conservative Trends in Agency Ideal Points

Figure \@ref(fig:f4repacs) reproduces the central empirical finding of @acsmapping2025: several agencies shift in a conservative direction over administrations, with the largest movement concentrated in a small subset of agencies. Most agencies cluster near the ideological center across all four administrations, while a minority drift substantially. This pattern is consistent with the author's argument that agency ideology is not static and responds to presidential incentives over time.

## Table \@ref(tab:t1repacs). Agencies with Largest Shifts in the Conservative Direction

Table \@ref(tab:t1repacs) identifies the agencies with the largest minimum-to-maximum range in estimated ideal points across administrations. The agencies at the top of this list represent the subset where executive pressure or personnel turnover most strongly reshaped ideological orientation. The replication reproduces the agency-level ordering and substantive interpretation from @acsmapping2025's Figure 4.

## Table \@ref(tab:t2repacs). Ideological Distance and Regulatory Productivity

Table \@ref(tab:t2repacs) replicates ACS Equation (3) across five outcome definitions. The coefficient on `distance` is negative across all columns, indicating that agencies whose ideology is further from the president produce fewer pages of final rules, fewer significant rules, and fewer regulatory-only rules. The negative binomial specification absorbs agency fixed effects and weights by the inverse variance of the ideal point estimates. The sign and magnitude patterns replicate the original Table 2 in @acsmapping2025.

The following subsections report results from the three methodological extensions. Each introduces a different outcome or estimation strategy to probe the scope of the ideological alignment mechanism identified by @acsmapping2025.

## Table \@ref(tab:extmodeltable). ACS Distance and Fazekas Donation/Contract Outcomes

Table \@ref(tab:extmodeltable) extends ACS Equation (3) by replacing regulatory activity outcomes with donation and contracting outcomes from @fazekasagency2023. The central parameter in each column is the coefficient on `distance`. The estimates are mixed across outcomes: `distance` is near zero or positive for donor contract share and donor contract count, but negative for log donation amount and total contracts. This pattern points to outcome sensitivity rather than a single uniform procurement effect.

## Figure \@ref(fig:extcoefplot). Distance Effect across Alternative Fazekas Outcomes

Figure \@ref(fig:extcoefplot) plots the `distance` coefficient from the extension models alongside $95\%$ confidence intervals. This visual comparison identifies which outcome definitions produce the most precise estimates and whether the direction of the distance effect is consistent across donor contract share, donor contract count, log donation amount, and total contracts. In this case, the figure shows that the sign is not uniform across outcomes, which reinforces the need to interpret the procurement extension as outcome-specific rather than as a single composite mechanism.

## Table \@ref(tab:ext1logit). Logistic Regression of Donation Probability on Ideological Distance

Table \@ref(tab:ext1logit) reports logistic regression estimates for the binary `donation_dummy` outcome. Column (1) includes only `distance` as a predictor; Column (2) adds an administration fixed effect. The coefficient on `distance` is positive in both specifications, which indicates that firms are more likely to donate when the agency's ideology is far from the president. The divide-by-4 rule (@gelman2025) converts the logit coefficient to an approximate upper bound on the marginal probability change per unit of ideological distance.

## Figure \@ref(fig:ext1logitplot). Predicted Donation Probability by Distance

Figure \@ref(fig:ext1logitplot) translates the logistic model into predicted donation probabilities across the observed range of `distance`, separately for Bush and Obama administrations. The upward-sloping curves indicate that firms are more likely to donate as agencies become more ideologically distant from the president. The administration comparison shows that this pattern appears in both presidencies, though at different baseline levels.

## Figure \@ref(fig:ext2ppc). Overdispersion Diagnostic

Figure \@ref(fig:ext2ppc) compares the distribution of zero-rates across 200 simulated datasets from Poisson and negative binomial models fitted to the ACS regulatory productivity data, alongside the observed zero-rate (dashed line). The Poisson model generates replicated zero-rates systematically below the observed proportion, indicating that its constraint $\text{sd}(y) = \sqrt{\text{E}(y)}$ is too restrictive for these data. The negative binomial model generates replications that surround the observed zero-rate, validating the overdispersion assumption in @acsmapping2025's Equation (3).

## Table \@ref(tab:ext3ordlogit). Ordered Logistic Regression of Agency Independence Tier

Table \@ref(tab:ext3ordlogit) reports the ordered logistic coefficient for `distance` on the four-tier agency independence scale, from most independent (indep. commission and IRC) to least independent (cabinet department). A negative $\hat{\beta}$ implies that agencies with greater ideological distance from the president tend to occupy more independent tiers, consistent with the institutional logic that structural independence reduces direct presidential alignment. A null result would indicate that ideological distance and formal independence are orthogonal properties of agencies in this sample.

## Table \@ref(tab:ext3aagencylevel). Agency-Level Ordered Logistic Regression of Independence Tier

Table \@ref(tab:ext3aagencylevel) re-estimates Extension 3 after collapsing the data to one row per agency and replacing administration-specific distance with mean ideological distance across administrations. This specification is closer to the institutional outcome because agency independence tier is fixed at the agency level. If the coefficient remains negative and is similar in magnitude to Table \@ref(tab:ext3ordlogit), the cross-administration result reflects a stable cross-agency pattern. If it attenuates toward zero, the earlier result is more likely to reflect within-agency repetition from repeated agency-administration observations rather than a genuine structural relationship between ideological distance and independence tier.


\newpage

# 5. Discussion

## Implications

This extension moves the dependent variable from regulatory output to procurement outcomes, testing whether the same ideological alignment mechanism identified in @acsmapping2025 also shapes distributive politics in federal contracting from @fazekasagency2023. The results suggest a more differentiated pattern than a simple extension of ACS: ideological distance clearly depresses regulatory output, but procurement outcomes respond differently depending on which outcome is measured. This broadens the contribution of the extension by showing that bureaucratic alignment may structure rule production and donor behavior through related but not identical channels.

The logistic and ordered logistic extensions add two dimensions to this picture. The logistic result indicates whether firms actively sort their donations across agencies as ideological distance changes, while the ordered logit tests whether ideological distance is itself a function of an agency's structural independence tier. Together, these three extensions separate the supply side (which agencies receive more regulatory output) from the demand side (which agencies attract more donation activity) of the political economy of bureaucratic alignment.

The agency-level follow-up to Extension 3 sharpens this interpretation by aligning the unit of analysis with the institutional outcome itself. If the relationship between ideological distance and agency independence persists after collapsing to one row per agency, the result is less likely to be an artifact of repeated agency-administration observations.

## Method-Theory Link

The theory in @acsmapping2025 holds that ideological distance between the president and an agency reduces the president's incentive to protect or prioritize that agency's output, thereby depressing regulatory productivity. My extension translates this mechanism into a procurement context using @fazekasagency2023. Here the logistic result points in the opposite direction from a simple alignment story: firms are more likely to donate to agencies that are ideologically farther from the president. One interpretation is that donations are used not only to reinforce alignment, but also to secure access where ideological distance creates greater uncertainty. That possibility matters more broadly for the literature because it suggests that political giving may serve both amplifying and compensatory functions in bureaucratic politics. The method follows directly from this theoretical chain: (1) logistic regression is the appropriate estimator when the outcome is a firm's binary donation decision, (2) negative binomial is appropriate for count outcomes where overdispersion is present, and (3) ordered logistic is appropriate when the outcome is a categorical independence hierarchy with more than two ordered levels.

## Limitations and Future Research

The principal limitation of this extension is the reliance on cross-dataset agency matching by name, which introduces measurement error when agency names differ across the @acsmapping2025 and @fazekasagency2023 data sources. To maximize panel coverage, I combine three matching strategies: strict matching against ACS agency identifiers, broad matching using Federal Register rulemaking records from 2004 to 2015, and fuzzy string matching for agencies unresolved by the first two approaches. Despite these efforts, the extension panel covers only the Bush and Obama administrations (2001-2016), as the @fazekasagency2023 dataset does not extend to the Trump period covered by @acsmapping2025. 
A more complete crosswalk using official agency code identifiers (RINs in the ACS data and USASpending codes in the Fazekas data) would further improve coverage and reduce matching error. 


More substantively, the current extension treats `distance` as time-varying at the administration level but does not interact it with presidential party or other institutional factors that might affect agencies' decision-making processes. @acsmapping2025 provides continuous ideal point estimates for 69 agencies across four administrations; a natural extension would interact these estimates with a partisan indicator or an agency ideology factor and examine whether ideological proximity amplifies donor-linked contracting (e.g., when the president and agency are aligned with the contributing party). I plan to explore this design in future work.


\newpage

## References
